### 2021-09-17 products/manuscript/Manuscript.Rmd
MADA Project part 1

---
title: "MADA Project Manuscript"
subtitle: ""
author: Dawson Dobash
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

# Summary/Abstract
_Write a summary of your project._

# Introduction

## General Background Information
COVID-19 hit the world by storm and is a possible ever-lasting change to the way society acts. It also had an impact on each countries funding towards health care. Actions within the past two years have been taken to help fight against this virus.

## Description of data and data source
The data I plan on using is from a mixture of locations. Two sources are from kaggle.com and the other is from the Organisation for Economic Co-operation and Development (OECD) website. My data consists of eight data sets with information over each country in the world (for those that data was collected for). Six of these data sets provide the same information just for different countries where each data set is a different region. There is a data set that includes all of the countries but I want to include a region indicator variable for each country. These data sets includes total covid-19 cases, active cases, total deaths, total recovered, total tests, and population size. This data was updated on kaggle as of Sept 16th. Another data set I am using is a vaccination data set for each country. This is also data as of Sept. 16th. Finally, the last data set is of each countries healthcare funding for the past 4-5 years. Some countries provide 2020 data but others only provide up to 2019.

I plan on combining the data sets into one with the variable combining them all is the country. If there are missing data within the variables used in analysis, I will exclude those countries. Ill have to change the raw data as one data set uses 3 letter abbreviations for countries while the others use the whole name of the country. There will be a lot of cleaning for this data just to make it as I want it.

I included 3 data sets out of the 8 within the processingscript for now because one shows the same information as the other data sets not included just in a different region of the world. This is just to show the data I will be using.

I plan on trying to find more information/data sets about some demographics data for the countries like type of government and an indicator for developed vs. developing countries.

Links to the locations I received the data are below:

https://data.oecd.org/healthres/health-spending.htm

https://www.kaggle.com/iamsouravbanerjee/covid19-dataset-world-and-continent-wise?select=Covid+Data+-+World.csv

https://www.kaggle.com/anandhuh/latest-worldwide-vaccine-data

## Questions/Hypotheses to be addressed
Since this data is cumulative and does not break down the amount of cases per month per country, I want to compare the recovery proportion to the death proportion and if this difference is based on variables such as the number of vaccines pushed out to society, the amount of funding the health care system has, the amount of tests taken. I also plan on adding more information as state in the description of data section. This will hopefully show if the amount of health care support really helps patients recover.

Some issues I can run into are: 
1) Developed countries might have more cases as they are in contact with the world more as travelers commonly travel to certain countries. Therefore could show an inverse relationship to what I expect with more developed countries have more recoveries. 
2) Unknown dates of when a country received vaccines.

# Methods and Results

_In most research papers, results and methods are separate. You can combine them here if you find it easier. You are also welcome to structure things such that those are separate sections._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Exploratory analysis

_Use a combination of text/tables/figures to explore and describe your data. You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. FIgures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data._

_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables. Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones._

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

Table \@ref(tab:summarytable) shows a table summarizing the data.

```{r summarytable,  echo=FALSE}
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable, caption = 'Data summary table.')
```

Figure \@ref(fig:resultfigure) shows a scatterplot figure produced by one of the R scripts.

```{r resultfigure,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/resultfigure.png")
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example table \@ref(tab:resulttable) shows a table summarizing a linear model fit.

```{r resulttable,  echo=FALSE}
resulttable=readRDS("../../results/resulttable.rds")
knitr::kable(resulttable, caption = 'Linear model fit table.')
```

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@Leek2015a] discusses types of analyses.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

# References

### 2021-09-17 code/analysis_code/analysisscript.R

@@ -8,6 +8,8 @@
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
+library(tidyverse)
+
 
 #path to data
 #note the use of the here() package and not absolute paths

### 2021-09-17 code/processing_code/processingscript.R

@@ -5,13 +5,17 @@
 #and saves it as Rds file in the processed_data folder
 
 #load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
+library(readr) #for loading CSV files
 library(dplyr) #for data processing
 library(here) #to set paths
+library(tidyverse)
 
 #path to data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
+data_location1 <- here::here("data","raw_data","Covid Data - Africa.csv")
+data_location2 <- here::here("data","raw_data","Healthcare_Funds.csv")
+data_location3 <- here::here("data","raw_data","Worldwide Vaccine Data.csv")
+
 
 #load data. 
 #note that for functions that come from specific packages (instead of base R)
@@ -19,37 +23,25 @@ data_location <- here::here("data","raw_data","exampledata.xlsx")
 #package::function() that's not required one could just call the function
 #specifying the package makes it clearer where the function "lives",
 #but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
+rawdata1 <- readr::read_csv(data_location1)
+rawdata2 <- readr::read_csv(data_location2)
+rawdata3 <- readr::read_csv(data_location3)
 
 #take a look at the data
-dplyr::glimpse(rawdata)
+dplyr::glimpse(rawdata1)
+dplyr::glimpse(rawdata2)
+dplyr::glimpse(rawdata3)
 
 #dataset is so small, we can print it to the screen.
 #that is often not possible.
-print(rawdata)
+print(rawdata1)
+print(rawdata2)
+print(rawdata3)
 
 # looks like we have measurements for height (in centimeters) and weight (in kilogram)
 
-# there are some problems with the data: 
-# There is an entry which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we need to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# We need to fix that too.
-# Then there is one person with a height of 6. 
-# that could be a typo, or someone mistakenly entered their height in feet.
-# Since we unfortunately don't know, we'll have to remove this person.
-# similarly, there is a person with weight of 7000, which is impossible,
-# and one person with missing weight.
[…]

### 2021-10-08 products/manuscript/Manuscript.Rmd
MADA Project part 2

@@
 Therefore could show an inverse relationship to what I expect with more developed countries have more recoveries.
-2) Unknown dates of when a country received vaccines.
+2) Unknown dates of when a country received vaccines and amount.
 
@@
 
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+## Data import and cleaning
+For the data importing and cleaning section, I first had to import 8 different data sets into R from multiple sources specified in the Description of data and data source.
+After this importation, The first thing to do before combing the data set was to create indicator variables for each country.
+After this I wanted to make sure that each variable was the right class as some of the numeric variables were character variables.
+This will help in making sure that each variable was able to be used in the right way.
 
-## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+I then was able to combine each continent data set into one world data set to make the importing go easier.
+After combining the data sets, I need to only include the variable for each data set needed for analysis and make sure that all na values were removed from each data set as I only want full information observations.
+However this would delete some countries that were large in population such as China and Phillipines as they did not have vaccine information.
+
+The last problem I ran into was that the country names for all datasets were different.
+One included a three letter code for countries, then another one used different abbreviations/names for countries.
+I had to use the countrynames package to help change into names I wanted.
+Since I could not change all of the countries names to the way I wanted it, I had to unfortunately go into the raw data csv files to change them manually.
+After all of this was good to go, I used the merge function to combine all of the data sets the way I wanted to.
+
+One disclosure I have for this is that one data set, health care funds, only had 48 countries when I cleaned the data.
+Therefore, there are a lot of na values for only those specific variables that came from the health care funds data set in the merged one.
+
+To look at a more descriptive explanation of all of the steps I took in data cleaning and importing, go to the processing script where it has comments for every line of code.
 
 ## Exploratory analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
-You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-FIgures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data._
-
-_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-Plots produced here can be scatterplots, boxplots, violinplots, etc.
-Tables can be simple 2x2 tables or larger ones._
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
 
@@
 
-Figure \@ref(fig:resultfigure) shows a scatterplot figure produced by one of the R scripts.
+Table \@ref(tab:table1) shows a table of the top 10 countries with the highest percentage of cases by population.
+We can see that there are a variety of locations and that Montenegro and Czech Republic are the highest 2 countries with the most cases by population.
 
-```{r resultfigure,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/resultfigure.png")
+```{r table1,  echo=FALSE}
+tab1=readRDS("../../results/table1.rds")
+knitr::kable(tab1)
+```
+
+Table \@ref(tab:table2) shows a table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate.
+For the bottom 5, there is a mixture of locations.
+We can also see that the recovery rate and death rate for the bottom 5 teams do not add up to 100 or close to it.
+This means there are a lot of active cases.
+
+```{r table2,  echo=FALSE}
+tab2=readRDS("../../results/table2.rds")
+knitr::kable(tab2)
+```
+
+Table \@ref(tab:table3) shows a table  of the top 5 and bottom 5 countries for death rate.
+We can see that Vanuatu only has a total of 4 cases but Yemen also has a significantly large rate then the next top country, Mexico.
+We can also see that Asia dominates the lowest death rates.
+
+```{r table3,  echo=FALSE}
+tab3=readRDS("../../results/table3.rds")
+knitr::kable(tab3)
+```
+
+Figure \@ref(fig:figure1) shows a scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.
+
+```{r figure1,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure1.png")
+```
+
+Figure \@ref(fig:figure2) shows a scatterplot figure of total cases by population filtered down to countries with less than 40 mllion population. We can see that the total cases are spread out.
+
+```{r figure2,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure2.png")
+```
+
+Figure \@ref(fig:figure3) shows a scatterplot figure of total recovered patients by total cases.
+We can see just like in figure 1 that the United States, India, and Brazil have a overwhelmingly more cases than other countries.
+Therefore in figure 4, we will explore the majority of the countries.
+
+```{r figure3,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure3.png")
+```
+
+Figure \@ref(fig:figure4) shows a scatterplot figure of the same plot as before just filtering countries with less than 20 million cases.
+We can see a linear relationship between the recovered to total cases.
+It looks like a slope of 1.
+
+```{r figure4,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure4.png")
+```
+
+Figure \@ref(fig:figure5) shows a histogram figure of the percent of fully vaccinated people.
+We can see that the histogram is right skewed where most of the countries have less than 25% of the population fully vaccinated.
+
+```{r figure5,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure5.png")
+```
+
+Figure \@ref(fig:figure6) shows a scatterplot figure of percent of people fully vaccinated by health care funding.
+We can see that it is a linear relationship with a positive correlation.
+Wecan also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.
+
+```{r figure6,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure6.png")
 ```
@@
 Then you load them here._
-
-Example table \@ref(tab:resulttable) shows a table summarizing a linear model fit.
-
-```{r resulttable,  echo=FALSE}
-resulttable=readRDS("../../results/resulttable.rds")
-knitr::kable(resulttable, caption = 'Linear model fit table.')
-```
 
@@
 
-This paper [@Leek2015a] discusses types of analyses.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
 # References

### 2021-10-08 code/analysis_code/analysisscript.R

@@ -21,51 +21,108 @@ mydata <- readRDS(data_location)
 ######################################
 #Data exploration/description
 ######################################
-#I'm using basic R commands here.
-#Lots of good packages exist to do more.
-#For instance check out the tableone or skimr packages
+#The most notable of countries not in the data set are China and Philippines as they did not have any vaccine numbers 
+#displayed within the source. 
 
-#summarize data 
-mysummary = summary(mydata)
+#This is just to look at the distributions of each numeric variable
+summarytable = summary(mydata)
 
-#look at summary
-print(mysummary)
+#For this scatter plot, I plotted each countries population compare to their total cases of COVID-19.
+#You can see that there are three countries that either has more population or total cases than others
+fig1 = mydata %>% ggplot(aes(x=Population, y=`Total Cases`, label=Country)) + geom_point() + 
+  geom_text(aes(label=ifelse(`Total Cases`>20000000, as.character(Country),'')), hjust=.3,vjust=1) + 
+            ggtitle("Figure 1: Scatterplot of Total Cases by Population for each Country \n Labeling some outliers")
 
-#do the same, but with a bit of trickery to get things into the 
-#shape of a data frame (for easier saving/showing in manuscript)
-summary_df = data.frame(do.call(cbind, lapply(mydata, summary)))
+#This scatter plot is in response to looking at only countries that has less than 40 million people. This is to
+#show the distribution of the other countries.
+fig2 = mydata %>% filter(Population <= 40000000) %>% ggplot(aes(x=Population, y=`Total Cases`)) + geom_point() + 
+  ggtitle("Figure 2: Scatterplot of Total Cases by Population for each Country \n under the Population of 40 million")
+
+#I wanted to see the top 10 Countries with the most cases proportionally to their population. I included the 
+#continent that are on just to see if there was any connection. We can see that there is a good mixture of countries 
+#within this top ten list
+tab1 = mydata %>% mutate(totcase_percent = (`Total Cases`/Population)*100) %>% arrange(desc(totcase_percent)) %>% 
+  select(Country,location, totcase_percent) %>% head(., 10)
+
+#This is a scatterplot of showing total cases by total recovered. We can see just like the population scatterplot
[…]

### 2021-10-08 code/processing_code/processingscript.R

@@ -5,52 +5,120 @@
 #and saves it as Rds file in the processed_data folder
 
 #load needed packages. make sure they are installed.
+
 library(readr) #for loading CSV files
 library(dplyr) #for data processing
 library(here) #to set paths
-library(tidyverse)
+library(tidyverse) #for cleaning and 'tidying' the data
+library(countrycode) #changes country codes to full names
 
 #path to data
 #note the use of the here() package and not absolute paths
+
 data_location1 <- here::here("data","raw_data","Covid Data - Africa.csv")
 data_location2 <- here::here("data","raw_data","Healthcare_Funds.csv")
 data_location3 <- here::here("data","raw_data","Worldwide Vaccine Data.csv")
+data_location4 <- here::here("data","raw_data","Covid Data - Asia.csv")
+data_location5 <- here::here("data","raw_data","Covid Data - Europe.csv")
+data_location6 <- here::here("data","raw_data","Covid Data - North America.csv")
+data_location7 <- here::here("data","raw_data","Covid Data - Oceania.csv")
+data_location8 <- here::here("data","raw_data","Covid Data - South America.csv")
+
+#load data from the data locations specified for each data set 
+
+africa <- readr::read_csv(data_location1)
+funds <- readr::read_csv(data_location2)
+vaccine <- readr::read_csv(data_location3)
+asia <- readr::read_csv(data_location4)
+europe <- readr::read_csv(data_location5)
+n_america <- readr::read_csv(data_location6)
+oceania <- readr::read_csv(data_location7)
+s_america <- readr::read_csv(data_location8)
 
 
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata1 <- readr::read_csv(data_location1)
-rawdata2 <- readr::read_csv(data_location2)
-rawdata3 <- readr::read_csv(data_location3)
+#take a look at the data.I only included looking at the africa data set because the other continent data sets have the
+#exact layout just different countries
 
-#take a look at the data
-dplyr::glimpse(rawdata1)
-dplyr::glimpse(rawdata2)
-dplyr::glimpse(rawdata3)
+dplyr::glimpse(africa)
+dplyr::glimpse(funds)
+dplyr::glimpse(vaccine)
 
[…]

### 2021-10-29 products/manuscript/Manuscript.Rmd
MADA Project part 3-
Cleaned up the repo and did some simple analysis

@@
 
-I plan on trying to find more information/data sets about some demographics data for the countries like type of government and an indicator for developed vs. developing countries.
-
 Links to the locations I received the data are below:
@@
 ## Questions/Hypotheses to be addressed
-Since this data is cumulative and does not break down the amount of cases per month per country, I want to compare the recovery proportion to the death proportion and if this difference is based on variables such as the number of vaccines pushed out to society, the amount of funding the health care system has, the amount of tests taken.
-I also plan on adding more information as state in the description of data section.
+1) Since this data is cumulative and does not break down the amount of cases per month per country, I want to compare the recovery proportion to the death proportion and if this difference is based on variables such as the number of vaccines pushed out to society, the amount of funding the health care system has, the amount of tests taken.
 This will hopefully show if the amount of health care support really helps patients recover.
+
+2) I will also look at how the percent of COVID-19 cases for each country is affected by multiple predictors such as the ones addressed in the above paragraph.
 
@@
 
-Table \@ref(tab:summarytable) shows a table summarizing the data.
+Table 1 shows a table summarizing the data.
 
@@
 
-Table \@ref(tab:table1) shows a table of the top 10 countries with the highest percentage of cases by population.
+Table 2 shows a table of the top 10 countries with the highest percentage of cases by population.
 We can see that there are a variety of locations and that Montenegro and Czech Republic are the highest 2 countries with the most cases by population.
@@
 
-Table \@ref(tab:table2) shows a table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate.
+Table 3 shows a table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate.
 For the bottom 5, there is a mixture of locations.
@@
 
-Table \@ref(tab:table3) shows a table  of the top 5 and bottom 5 countries for death rate.
+Table 4 shows a table  of the top 5 and bottom 5 countries for death rate.
 We can see that Vanuatu only has a total of 4 cases but Yemen also has a significantly large rate then the next top country, Mexico.
@@
 
-Figure \@ref(fig:figure1) shows a scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.
+Figure 1 shows a scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.
 
@@
 
-Figure \@ref(fig:figure2) shows a scatterplot figure of total cases by population filtered down to countries with less than 40 mllion population. We can see that the total cases are spread out.
+Figure 2 shows a scatterplot figure of total cases by population filtered down to countries with less than 40 million population. We can see that the total cases are spread out.
 
@@
 
-Figure \@ref(fig:figure3) shows a scatterplot figure of total recovered patients by total cases.
+Figure 3 shows a scatterplot figure of total recovered patients by total cases.
 We can see just like in figure 1 that the United States, India, and Brazil have a overwhelmingly more cases than other countries.
@@
 
-Figure \@ref(fig:figure4) shows a scatterplot figure of the same plot as before just filtering countries with less than 20 million cases.
+Figure 4 shows a scatterplot figure of the same plot as before just filtering countries with less than 20 million cases.
 We can see a linear relationship between the recovered to total cases.
@@
 
-Figure \@ref(fig:figure5) shows a histogram figure of the percent of fully vaccinated people.
+Figure 5 shows a histogram figure of the percent of fully vaccinated people.
 We can see that the histogram is right skewed where most of the countries have less than 25% of the population fully vaccinated.
@@
 
-Figure \@ref(fig:figure6) shows a scatterplot figure of percent of people fully vaccinated by health care funding.
+Figure 6 shows a scatterplot figure of percent of people fully vaccinated by health care funding.
 We can see that it is a linear relationship with a positive correlation.
-Wecan also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.
+We can also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.
 
@@
 
+Figure 7 shows a histogram figure that describes the difference in proportions between those who recovered versus the proportion of those who died.
+From the histogram we can see that this difference is right skewed as most countries have a much hagher recovery rate with only a some having more of a lower difference meaning lower recovery rates.
+
+```{r figure7,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure7.png")
+```
+
 ## Full analysis
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
+### Difference in proportions of recovery and deaths
+
+For my first question, we are looking at the difference in proportions of those who have recovered versus those how died.
+We can see in figure 7 that this data is fairly skewed.
+Therefore, I will do a logit transformation of the outcome to make it more normal as shown in figure 8.
+
+```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure8.png")
+```
+
+Before I decided on the logit transformation, I first decided to try this difference in proportions outcome I made within a binomial model and then a quasibinomial using the values of total cases as my weight to see how the deviance and dispersion values are.
+After running two simple models, the results were showing very high values for these statistics.
+Therefore this is how I decided on using the logit transformation.
+
+Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of each of the models within Tables 5-8.
+
+Table 5 is showing statistics for the predictor vaccine percentage.
+```{r table5,  echo=FALSE}
+tab5=readRDS("../../results/tablevacc1.rds")
+knitr::kable(tab5)
+```
+
+Table 6 is showing statistics for the predictor health care funds.
+```{r table6,  echo=FALSE}
+tab6=readRDS("../../results/tablehc1.rds")
+knitr::kable(tab6)
+```
+
+Table 7 is showing statistics for the predictor number of tests per person.
+```{r table7,  echo=FALSE}
+tab7=readRDS("../../results/tabletests1.rds")
+knitr::kable(tab7)
+```
+
+Table 8 is showing statistics for the predictor location which specifies the continent for each country.
+```{r table8,  echo=FALSE}
+tab8=readRDS("../../results/tableloc1.rds")
+knitr::kable(tab8)
+```
+
+Looking at all of these tables, we can see that the R^2 values are very low.
+Therefore none of these predictors shows any association with the logit transformation outcome.
+Since not even one predictor is good for this outcome, I am going to spot my analysis with this outcome and go to my second question which uses the percent of cases within each country.
+
+### Percent of cases for each country
 

### 2021-10-29 code/analysis_code/analysisscript.R

@@ -9,6 +9,9 @@ library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
 library(tidyverse)
+library(AER)
+library(tidymodels)
+
 
 
 #path to data
@@ -87,6 +90,12 @@ fig6 = mydata %>% ggplot(aes(x=Value, y=`% of population fully vaccinated`, labe
   ggtitle("Figure 6: Scatterplot of Percent of People Fully Vaccinated by Health Care Funds \n (by Country)") +
   xlab("Health Care Funds")
 
+#This is a histogram of the difference in proportions of those who are recovered and who died with each denominator as
+#number of cases.
+fig7 = mydata %>% ggplot(aes(x = prop_diff_recov_vs_death )) + geom_histogram(binwidth = 0.05) + 
+  ggtitle("Figure 7: Histogram of Difference in proportions \n of those who recovered versus those who died \n (by Country)")
+fig7
+
 #save data frame table to file for later use in manuscript
 summarytable_file = here("results", "summarytable.rds")
 saveRDS(summarytable, file = summarytable_file)
@@ -101,7 +110,7 @@ tab3_file = here("results", "table3.rds")
 saveRDS(tab3, file = tab3_file)
 
 
-#save figure
+#save figures
 fig1_file = here("results","figure1.png")
 ggsave(filename = fig1_file, plot=fig1) 
 
@@ -120,9 +129,106 @@ ggsave(filename = fig5_file, plot=fig5)
 fig6_file = here("results","figure6.png")
 ggsave(filename = fig6_file, plot=fig6) 
 
+fig7_file = here("results","figure7.png")
+ggsave(filename = fig7_file, plot=fig7)
+
 ######################################
 #Data fitting/statistical analysis
 ######################################
 
+#Models 1 and 11 are looking at making models using binomial and quasibinomial as their family types. Included were summary tables
+#for each that show model information
+model1 = glm(prop_diff_recov_vs_death ~ prop_vacc, weights = `Total Cases`, data = mydata, family = "binomial")
+summary(model1)
 
-  
+model11 = glm(prop_diff_recov_vs_death ~ prop_vacc, weights = `Total Cases`, data = mydata, family = "quasibinomial")
+summary(model11)
+
+
+
+
+
+
[…]

### 2021-10-29 code/processing_code/processingscript.R

@@ -110,8 +110,16 @@ merged_data = merge(clean_world, clean_vaccine, by="Country")
 #Since this data set is only 48 countries, it want to merge it into the combined data set but also including all 
 #observations from the combined data set. I am also removing the variable LOCATION as it is a redundant variable now
 
-processeddata = merge(merged_data, clean_funds2, by="Country", all.x = T)
-processeddata = select(processeddata, -LOCATION)
+merged_data2 = merge(merged_data, clean_funds2, by="Country", all.x = T)
+merged_data2 = select(merged_data2, -LOCATION)
+
+#Need to create some more variables such as proportions
+processeddata <- merged_data2 %>% mutate(prop_death = `Total Deaths`/`Total Cases`,
+                                         prop_recov = `Total Recovered`/`Total Cases`,
+                                         pct_cases = (`Total Cases`/Population)*100,
+                                         test_per_person = `Total Tests`/Population,
+                                         prop_diff_recov_vs_death = prop_recov - prop_death,
+                                         location = as.factor(location)) %>% filter(`Total Cases` >= 1000)
 
 #Looking at the complete data set
 

### 2021-10-29 products/manuscript/Manuscript.Rmd
Added some more analysis to Part 3

@@
 Figure 7 shows a histogram figure that describes the difference in proportions between those who recovered versus the proportion of those who died.
-From the histogram we can see that this difference is right skewed as most countries have a much hagher recovery rate with only a some having more of a lower difference meaning lower recovery rates.
+From the histogram we can see that this difference is left skewed as most countries have a much hagher recovery rate with only a some having more of a lower difference meaning lower recovery rates.
 
@@
 knitr::include_graphics("../../results/figure7.png")
+```
+
+Figure 8 shows a histogram figure of the percentage of cases per population. We can see that this data is highly skewed to the right.
+
+```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure8.png")
 ```
@@
 We can see in figure 7 that this data is fairly skewed.
-Therefore, I will do a logit transformation of the outcome to make it more normal as shown in figure 8.
+Therefore, I will do a logit transformation of the outcome to make it more normal as shown in figure 9.
 
-```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure8.png")
+```{r figure9,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/figure9.png")
 ```
@@
 
+We can see with looking at figure 8 that the percentage of cases are extremely right skewed.
+For now, I will proceed with doing some simple linear regression.
+Below shows model statistics for each model.
+
+Table 9 is showing statistics for the predictor vaccine percentage.
+```{r table9,  echo=FALSE}
+tab9=readRDS("../../results/tablevacc2.rds")
+knitr::kable(tab9)
+```
+
+Table 10 is showing statistics for the predictor health care funds.
+```{r table10,  echo=FALSE}
+tab10=readRDS("../../results/tablehc2.rds")
+knitr::kable(tab10)
+```
+
+Table 11 is showing statistics for the predictor number of tests per person.
+```{r table11,  echo=FALSE}
+tab11=readRDS("../../results/tabletests2.rds")
+knitr::kable(tab11)
+```
+
+Table 12 is showing statistics for the predictor location which specifies the continent for each country.
+```{r table12,  echo=FALSE}
+tab12=readRDS("../../results/tableloc2.rds")
+knitr::kable(tab12)
+```
+
+From these tables, we can see that the R^2's for these models are still relatively low but a simple linear regression of using percent vaccination or the location of each country have R^2's of around 0.35 which is an improvement from the linear regression models from using logit transformation of difference in proportions as the outcome.
+
 # Discussion

### 2021-10-29 code/analysis_code/analysisscript.R

@@ -96,6 +96,12 @@ fig7 = mydata %>% ggplot(aes(x = prop_diff_recov_vs_death )) + geom_histogram(bi
   ggtitle("Figure 7: Histogram of Difference in proportions \n of those who recovered versus those who died \n (by Country)")
 fig7
 
+#Looking at the distribution using a histogram
+fig8 = mydata %>% ggplot(aes(x = pct_cases)) + geom_histogram(binwidth = 2) + 
+  ggtitle("Figure 8: Histogram of Percentage of cases\n (by Country)")
+fig8
+
+
 #save data frame table to file for later use in manuscript
 summarytable_file = here("results", "summarytable.rds")
 saveRDS(summarytable, file = summarytable_file)
@@ -132,6 +138,10 @@ ggsave(filename = fig6_file, plot=fig6)
 fig7_file = here("results","figure7.png")
 ggsave(filename = fig7_file, plot=fig7)
 
+fig8_file = here("results","figure8.png")
+ggsave(filename = fig8_file, plot=fig8)
+
+
 ######################################
 #Data fitting/statistical analysis
 ######################################
@@ -160,13 +170,13 @@ mydata$logit_prop_diff = logit(mydata$prop_diff_recov_vs_death)
 
 #This is a histogram of the logit difference in proportions of those who are recovered and who died with each denominator as
 #number of cases.
-fig8 = mydata %>% ggplot(aes(x = logit_prop_diff )) + geom_histogram(binwidth = 0.5) + 
-  ggtitle("Figure 8: Histogram of logit transformation of difference in proportions \n of those who recovered versus those who died \n (by Country)")
-fig8
+fig9 = mydata %>% ggplot(aes(x = logit_prop_diff )) + geom_histogram(binwidth = 0.5) + 
+  ggtitle("Figure 9: Histogram of logit transformation of difference in proportions \n of those who recovered versus those who died \n (by Country)")
+fig9
 
 #Saving figure * to a location for future references
-fig8_file = here("results","figure8.png")
-ggsave(filename = fig8_file, plot=fig8)
+fig9_file = here("results","figure9.png")
+ggsave(filename = fig9_file, plot=fig9)
 
 
[…]

### 2021-10-29 products/manuscript/Manuscript.Rmd
Added titles to each figure for part 3

--- 2021-10-29 products/manuscript/Manuscript.Rmd
@@
 
+We can see with looking at figure 8 that the percentage of cases are extremely right skewed.
+For now, I will proceed with doing some simple linear regression.
+Below shows model statistics for each model and a visual showing the relationship between the two variables.
+
+Figure 10 is showing statistics for the predictor vaccine percentage.
+```{r vaccine pct,  echo=FALSE}
+knitr::include_graphics("../../results/figure10.png")
+
+tab9=readRDS("../../results/tablevacc2.rds")
+knitr::kable(tab9)
+```
+
+Figure 11 is showing statistics for the predictor health care funds.
+```{r health care,  echo=FALSE}
+knitr::include_graphics("../../results/figure11.png")
+
+tab10=readRDS("../../results/tablehc2.rds")
+knitr::kable(tab10)
+```
+
+Figure 12 is showing statistics for the predictor number of tests per person.
+```{r Tests,  echo=FALSE}
+knitr::include_graphics("../../results/figure12.png")
+
+tab11=readRDS("../../results/tabletests2.rds")
+knitr::kable(tab11)
+```
+
+Figure 13 is showing statistics for the predictor location which specifies the continent for each country.
+```{r Location,  echo=FALSE}
+knitr::include_graphics("../../results/figure13.png")
+
+tab12=readRDS("../../results/tableloc2.rds")
+knitr::kable(tab12)
+```
+
+From these visuals, we can see that the R^2's for these models are still relatively low but a simple linear regression of using percent vaccination or the location of each country have R^2's of around 0.35 which is an improvement from the linear regression models from using logit transformation of difference in proportions as the outcome.
+
 # Discussion

### 2021-10-29 code/analysis_code/analysisscript.R

@@ -282,15 +282,37 @@ saveRDS(loc_stats2, file = tabloc2_file)
 
 #Visualization of the simple linear regression models
 #Simple scatterplots of these simple linear regressions
-vacc_scatterplot <- mydata %>% ggplot(aes(x=`% of population vaccinated`, y = pct_cases)) + geom_point()
+vacc_scatterplot <- mydata %>% ggplot(aes(x=`% of population vaccinated`, y = pct_cases)) + geom_point() + 
+  ggtitle("Figure 10: Scatterplot of Percent of People Fully Vaccinated and Percentage of Cases \n (by Country)") 
 vacc_scatterplot
 
-hc_fund_scatterplot <- mydata %>% ggplot(aes(x=Value, y = pct_cases)) + geom_point()
+hc_fund_scatterplot <- mydata %>% ggplot(aes(x=Value, y = pct_cases)) + geom_point() + 
+  ggtitle("Figure 11: Scatterplot of Healthcare funds and Percentage of Cases \n (by Country)") +
+  xlab("Health Care Funds")
 hc_fund_scatterplot
 
 tests_scatterplot <- mydata %>% ggplot(aes(x=test_per_person, y = pct_cases)) + geom_point()
 tests_scatterplot
 
+tests_scatterplot2 <- mydata %>% filter(test_per_person < 2) %>% ggplot(aes(x=test_per_person, y = pct_cases)) + geom_point()  + 
+  ggtitle("Figure 12: Scatterplot of COVID-19 tests per person and Percentage of Cases \n (by Country)") 
+tests_scatterplot2
+
 #Simple boxplot of difference in proportions by continent
-loc_boxplot <- mydata %>% ggplot(aes(x=location, y = pct_cases)) + geom_boxplot()
+loc_boxplot <- mydata %>% ggplot(aes(x=location, y = pct_cases)) + geom_boxplot()  + 
+  ggtitle("Figure 13: Histogram of the percent of cases by location \n (for each Country)") +
+  xlab("Continent")
 loc_boxplot
+
+#Save figures to result folder
+fig10_file = here("results","figure10.png")
+ggsave(filename = fig10_file, plot=vacc_scatterplot)
+
+fig11_file = here("results","figure11.png")
+ggsave(filename = fig11_file, plot=hc_fund_scatterplot)
+
+fig12_file = here("results","figure12.png")
+ggsave(filename = fig12_file, plot=tests_scatterplot2)
+
[…]

### 2021-10-29 products/manuscript/Manuscript.Rmd
Small Changes to the manuscript (Part 3)

@@
 Therefore none of these predictors shows any association with the logit transformation outcome.
-Since not even one predictor is good for this outcome, I am going to spot my analysis with this outcome and go to my second question which uses the percent of cases within each country.
+Since not even one predictor is good for this outcome, I am going to stop my analysis with this outcome and go to my second question which uses the percent of cases within each country.
+However, the p-values for all models are significant which does not makes sense when the r-squared values are really low.
 
@@
 From these visuals, we can see that the R^2's for these models are still relatively low but a simple linear regression of using percent vaccination or the location of each country have R^2's of around 0.35 which is an improvement from the linear regression models from using logit transformation of difference in proportions as the outcome.
+We can also see that for Figure 11 the scatterplot does not show that much association compared to the Figures 10, 12, and 13.
+We can see in Figure 13, the boxplots have some significant differences such as Africa compared to the other continents like Europe, South America, and North America.
 

### 2021-10-29 code/analysis_code/analysisscript.R

@@ -316,3 +316,16 @@ ggsave(filename = fig12_file, plot=tests_scatterplot2)
 
 fig13_file = here("results","figure13.png")
 ggsave(filename = fig13_file, plot=loc_boxplot)
+
+
+
+
+
+
+#Multivariable regression
+mult_reg_rec1 = recipe(pct_cases ~ `% of population vaccinated` + location, data=mydata)
+
+mult_reg_wrkflow1 <- workflow() %>% add_model(lm_mod) %>% add_recipe(mult_reg_rec1)
+mult_reg_fit1 <- mult_reg_wrkflow1 %>% fit(data = mydata)
+mult_reg_fit1 %>% extract_fit_parsnip() %>% tidy()
+mult_reg_stats = glance(mult_reg_fit1)

### 2021-11-12 products/manuscript/Manuscript.Rmd
MADA Project part 4

--- 2021-10-29 products/manuscript/Manuscript.Rmd
@@
 Actions within the past two years have been taken to help fight against this virus.
+Although COVID-19 has been here for over two year, there is still problems with under staffing in hospitals and not enough beds at some hospitals.
+Recovery rates are still high in the United States but what about the whole world.
+Other countries are still getting hit hard by this virus.
+In terms of vaccination output, there has been recent back and forth movements on people getting the vaccination.
+Although in developed countries this happens but what about the less developed countries and regions such as Africa.
+While people in the United States are fighting about whether the vaccine should be given or not, the people in these other countries might have a limited supply.
+
+Looking at the difference in recovered patients and those who have died will be beneficial in determining if more care is needed in certain countries.
+I also think the percentage of people vaccinated is also affected and can lead to important results that might show we need more either distribution or requirements of the vaccines.
 
 ## Description of data and data source
-The data I plan on using is from a mixture of locations.
-Two sources are from kaggle.com and the other is from the Organisation for Economic Co-operation and Development (OECD) website.
-My data consists of eight data sets with information over each country in the world (for those that data was collected for).
+The data I am using is from a mixture of locations.
+Two sources are from kaggle.com, one from the world bank database, one created by myself based on information from wikipedia, and the other is from the Organisation for Economic Co-operation and Development (OECD) website.
+My data consists of ten data sets with information over each country in the world (for those that data was collected for).
 Six of these data sets provide the same information just for different countries where each data set is a different region.
@@
 16th.
-Finally, the last data set is of each countries healthcare funding for the past 4-5 years.
-Some countries provide 2020 data but others only provide up to 2019.
+The other data sets are providing single stats for each country that was not included in the larger data sets but I still wanted to use.
+These data sets include the following statistics/demographics: healthcare funding for 2019-2020, vaccination numbers/percentages, gdp per capita, and government type.
 
@@
 If there are missing data within the variables used in analysis, I will exclude those countries.
-Ill have to change the raw data as one data set uses 3 letter abbreviations for countries while the others use the whole name of the country.
-There will be a lot of cleaning for this data just to make it as I want it.
-
-I included 3 data sets out of the 8 within the processingscript for now because one shows the same information as the other data sets not included just in a different region of the world.
-This is just to show the data I will be using.
+There will be a lot of cleaning for this data just to make it as I want it and work.
 
@@
 
+https://data.worldbank.org/indicator/NY.GDP.PCAP.CD
+
+https://en.wikipedia.org/wiki/List_of_countries_by_system_of_government
+
 ## Questions/Hypotheses to be addressed
@@
 
-2) I will also look at how the percent of COVID-19 cases for each country is affected by multiple predictors such as the ones addressed in the above paragraph.
+2) I will also look at how the percent of people vaccinated for each country is affected by multiple predictors.
+I am planning on seeing that more developed countries based on GDP per capita will have a more vaccinated population.
 
-Some issues I can run into are: 
-1) Developed countries might have more cases as they are in contact with the world more as travelers commonly travel to certain countries.
-Therefore could show an inverse relationship to what I expect with more developed countries have more recoveries.
-2) Unknown dates of when a country received vaccines and amount.
+Some confounders that I could run into are: 
+1) Developed countries might have more cases as more travelers visit these countries and it could have an effect on recovery numbers.
+
+2) Unknown dates of when a country received vaccines and amount. This could skew the amount of vaccinations a country receives.
 
@@
 
-_In most research papers, results and methods are separate. You can combine them here if you find it easier. You are also welcome to structure things such that those are separate sections._
+## Data import and cleaning
+For data importing and cleaning, I first had to import ten different data sets into R from multiple sources specified in the Description of data and data source.
+After this importation, The first thing to do before combining the data set was to create indicator variables for each country.
+After this, I wanted to make sure that each variable was the right class as some of the numeric variables were character variables.
+This will help in making sure that each variable was able to be used in the right way.
+I then was able to combine each continent data set into one world data set to make the importing go easier.
+After combining the continent data sets, I need to only include the variables from the world data set needed for analysis and make sure that all missing values were removed as I only want full information observations.
+However this would delete some countries that were large in population such as China and the Philippines as they did not have vaccine information.
 
-## Data import and cleaning
-For the data importing and cleaning section, I first had to import 8 different data sets into R from multiple sources specified in the Description of data and data source.
-After this importation, The first thing to do before combing the data set was to create indicator variables for each country.
-After this I wanted to make sure that each variable was the right class as some of the numeric variables were character variables.
-This will help in making sure that each variable was able to be used in the right way.
-
-I then was able to combine each continent data set into one world data set to make the importing go easier.
-After combining the data sets, I need to only include the variable for each data set needed for analysis and make sure that all na values were removed from each data set as I only want full information observations.
-However this would delete some countries that were large in population such as China and Phillipines as they did not have vaccine information.
-
-The last problem I ran into was that the country names for all datasets were different.
-One included a three letter code for countries, then another one used different abbreviations/names for countries.
-I had to use the countrynames package to help change into names I wanted.
-Since I could not change all of the countries names to the way I wanted it, I had to unfortunately go into the raw data csv files to change them manually.
+The next part was looking at all of the other datasets which include the government type, gdp per capita, health care funds, and vaccine datasets.
+To combine these with the world data set, there was a huge problem that goes with using data sets from different sources.
+This problem is how to combine the data sets with a specific variable.
+Fortunately, each data set had a country variable to combine them with.
+However, the country names for all datasets were different.
+One included a three letter code for countries, then another ones used different abbreviations/names for countries.
+I had to use the countrynames package to help change into names I wanted for the three-letter code.
+Since I could not change all of the countries names to the way I wanted it, I had to unfortunately go into the raw data csv files to change them manually for some of the continent specific data sets.
+I also used `ifelse` statements to change the names for the gdp per capita and health care funds data sets.
+I was able to do these in r because there were not as many country names to change compared to the other data sets I had to manually change.
+A comment within the proccessingscript.R tell of the countries that needed manual changes.
 After all of this was good to go, I used the merge function to combine all of the data sets the way I wanted to.
 
+I created some variables from the pre-existing ones that I will use within my analysis. Some of these changes were, for example, creating proportions from percentages and combining two variables.
+
 One disclosure I have for this is that one data set, health care funds, only had 48 countries when I cleaned the data.
-Therefore, there are a lot of na values for only those specific variables that came from the health care funds data set in the merged one.
+Therefore, there are a lot of missing values for only the specific variables that came from the health care funds data set in the merged one.
 
 To look at a more descriptive explanation of all of the steps I took in data cleaning and importing, go to the processing script where it has comments for every line of code.
-
-## Exploratory analysis
-
-Table 1 shows a table summarizing the data.
-
-```{r summarytable,  echo=FALSE}
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable, caption = 'Data summary table.')
-```
-
-Table 2 shows a table of the top 10 countries with the highest percentage of cases by population.
-We can see that there are a variety of locations and that Montenegro and Czech Republic are the highest 2 countries with the most cases by population.
-
-```{r table1,  echo=FALSE}
-tab1=readRDS("../../results/table1.rds")
-knitr::kable(tab1)
-```
-
-Table 3 shows a table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate.
-For the bottom 5, there is a mixture of locations.
-We can also see that the recovery rate and death rate for the bottom 5 teams do not add up to 100 or close to it.
-This means there are a lot of active cases.
-
-```{r table2,  echo=FALSE}
-tab2=readRDS("../../results/table2.rds")
-knitr::kable(tab2)
-```
-
-Table 4 shows a table  of the top 5 and bottom 5 countries for death rate.
-We can see that Vanuatu only has a total of 4 cases but Yemen also has a significantly large rate then the next top country, Mexico.
-We can also see that Asia dominates the lowest death rates.
-
-```{r table3,  echo=FALSE}
-tab3=readRDS("../../results/table3.rds")
-knitr::kable(tab3)
-```
-
-Figure 1 shows a scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.
-
-```{r figure1,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure1.png")
-```
-
-Figure 2 shows a scatterplot figure of total cases by population filtered down to countries with less than 40 million population. We can see that the total cases are spread out.
-
-```{r figure2,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure2.png")
-```
-
-Figure 3 shows a scatterplot figure of total recovered patients by total cases.
-We can see just like in figure 1 that the United States, India, and Brazil have a overwhelmingly more cases than other countries.
-Therefore in figure 4, we will explore the majority of the countries.
-
-```{r figure3,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure3.png")
-```
-
-Figure 4 shows a scatterplot figure of the same plot as before just filtering countries with less than 20 million cases.
-We can see a linear relationship between the recovered to total cases.
-It looks like a slope of 1.
-
-```{r figure4,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure4.png")
-```
-
-Figure 5 shows a histogram figure of the percent of fully vaccinated people.
-We can see that the histogram is right skewed where most of the countries have less than 25% of the population fully vaccinated.
-
-```{r figure5,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure5.png")
-```
-
-Figure 6 shows a scatterplot figure of percent of people fully vaccinated by health care funding.
-We can see that it is a linear relationship with a positive correlation.
-We can also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.
-
-```{r figure6,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure6.png")
-```
-
-Figure 7 shows a histogram figure that describes the difference in proportions between those who recovered versus the proportion of those who died.
-From the histogram we can see that this difference is left skewed as most countries have a much hagher recovery rate with only a some having more of a lower difference meaning lower recovery rates.
-
-```{r figure7,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure7.png")
-```
-
-Figure 8 shows a histogram figure of the percentage of cases per population. We can see that this data is highly skewed to the right.
-
-```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure8.png")
-```
 
@@
 
-For my first question, we are looking at the difference in proportions of those who have recovered versus those how died.
-We can see in figure 7 that this data is fairly skewed.
-Therefore, I will do a logit transformation of the outcome to make it more normal as shown in figure 9.
+Table 1.1 displays the top five and bottom five death rates.
+Also within this, it shows their corresponding recovery rates as well.
+We can see that the bottom five death rates are from Asia but also the highest, Yemen, is also within the Asia continent as well.
 
-```{r figure9,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure9.png")
+Table 1.1
+```{r table1,  echo=FALSE}
+tab1=readRDS("../../results/table3.rds")
+knitr::kable(tab1)
+```
+
+Figure 1.1 shows a histogram figure that describes the difference in proportions between those who recovered versus the proportion of those who died.
+From the histogram we can see that this difference is left skewed as most countries have a much higher recovery rate with only a some having more of a lower difference meaning lower recovery rates.
+This is very worrisome for the analysis as most of the countries are over 0.80 in the difference.
+
+```{r figure 1.1,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/prop_diff_recov_died.png")
+```
+
+We can see in figure 1.1 that this data is fairly skewed. Therefore, I will do a logit transformation of the outcome to make it more normal as shown in figure 1.2.
+
+```{r figure 1.2,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/log_prop_diff_hist.png")
 ```
@@
 
-Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of each of the models within Tables 5-8.
+Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of each of the models within Tables 1.2-1.5.
 
-Table 5 is showing statistics for the predictor vaccine percentage.
-```{r table5,  echo=FALSE}
-tab5=readRDS("../../results/tablevacc1.rds")
-knitr::kable(tab5)
+Table 1.2 is showing statistics for the predictor vaccine percentage.
+
+Table 1.2
+```{r table 1.2,  echo=FALSE}
+tab2=readRDS("../../results/tablevacc1.rds")
+knitr::kable(tab2)
 ```
 
-Table 6 is showing statistics for the predictor health care funds.
-```{r table6,  echo=FALSE}
-tab6=readRDS("../../results/tablehc1.rds")
-knitr::kable(tab6)
+Table 1.3 is showing statistics for the predictor health care funds.
+
+Table 1.3
+```{r table 1.3,  echo=FALSE}
+tab3=readRDS("../../results/tablehc1.rds")
+knitr::kable(tab3)
 ```
 
-Table 7 is showing statistics for the predictor number of tests per person.
-```{r table7,  echo=FALSE}
-tab7=readRDS("../../results/tabletests1.rds")
-knitr::kable(tab7)
+Table 1.4 is showing statistics for the predictor number of tests per person.
+
+Table 1.4
+```{r table 1.4,  echo=FALSE}
+tab4=readRDS("../../results/tabletests1.rds")
+knitr::kable(tab4)
 ```
 
-Table 8 is showing statistics for the predictor location which specifies the continent for each country.
-```{r table8,  echo=FALSE}
-tab8=readRDS("../../results/tableloc1.rds")
-knitr::kable(tab8)
+Table 1.5 is showing statistics for the predictor location which specifies the continent for each country.
+
+Table 1.5
+```{r table 1.5,  echo=FALSE}
+tab5=readRDS("../../results/tableloc1.rds")
+knitr::kable(tab5)
 ```
@@
 Therefore none of these predictors shows any association with the logit transformation outcome.
-Since not even one predictor is good for this outcome, I am going to spot my analysis with this outcome and go to my second question which uses the percent of cases within each country.
+Since not even one predictor is good for this outcome, I am going to stop my analysis with this outcome and go to my second question which uses the percent of cases within each country.
 
-### Percent of cases for each country
+### Percent of people vaccinated for each country
 
-We can see with looking at figure 8 that the percentage of cases are extremely right skewed.
-For now, I will proceed with doing some simple linear regression.
-Below shows model statistics for each model.
+#### Methods
 
-Table 9 is showing statistics for the predictor vaccine percentage.
-```{r table9,  echo=FALSE}
-tab9=readRDS("../../results/tablevacc2.rds")
-knitr::kable(tab9)
+For my main analysis, I will be using the logit transformation of the proportion of people vaccinated within each country as my main outcome.
+Since I am using the logit transformation, I am creating another data set that is a subset of the overall data set that only includes the variables I will be using for my analysis.
+The predictor variables I will be testing out on the main outcome are: percent of cases within a population, continent the country is on, number of tests per person, gdp per capita, proportion of those who died, government type, and health care funds.
+These will create my full model.
+
+The first part of my main analysis will be using test/training data and finding the RMSE for both the full model and the null model which only include a intercept.
+This is just to see if using a full model is better than the null model.
+
+After comparing the two RMSE, I will be going into more detail on model fitting and tuning.
+I will be applying my full model a machine learning model, the LASSO model.
+This will help find the best model I can use for the predictors in predicting the logit transformation of the proportion of people vaccinated.
+I will be applying this lasso model to both my test and training data created before to make sure the model fits both data sets similarly so show improvement on over-fitting.
+
+#### Exploratory Analysis
+
+Table 2.1 displays percentage of people vaccinated within the top five and bottom five countries.
+Just by the look of the countries displayed, the less developed countries tend to not have a lot of people vaccinated.
+It is honestly kind of scary to see some countries are not vaccinated at all or very low.
+
+Table 2.1
+```{r table 2.1,  echo=FALSE}
+tab21=readRDS("../../results/vacc_tab.rds")
+knitr::kable(tab21)
 ```
 
-Table 10 is showing statistics for the predictor health care funds.
-```{r table10,  echo=FALSE}
-tab10=readRDS("../../results/tablehc2.rds")
-knitr::kable(tab10)
+Figure 2.1 shows a histogram figure of the percent of vaccinated people. We can see that the histogram is right skewed with the spike around the percentages of 0%.
+
+```{r figure 2.1,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/vacc_hist.png")
 ```
 
-Table 11 is showing statistics for the predictor number of tests per person.
-```{r table11,  echo=FALSE}
-tab11=readRDS("../../results/tabletests2.rds")
-knitr::kable(tab11)
+Figure 2.2 shows that there is no trend between vaccinations and cases until you get out to over 10% of the population had cases of COVID-19.
+The vaccination numbers range a lot when there is a low perecentage of cases compared to the countries population.
+
+```{r figure 2.2,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/cases_vs_vacc.png")
 ```
 
-Table 12 is showing statistics for the predictor location which specifies the continent for each country.
-```{r table12,  echo=FALSE}
-tab12=readRDS("../../results/tableloc2.rds")
-knitr::kable(tab12)
+When looking at Figure 2.3, we can see that those countries who have large values of GDP per capita, their population is more vaccinated.
+The vaccination number, just like in Figure 2.2, varies when the GDP per capita is relatively low.
+
+```{r figure 2.3,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/gdp_vs_vacc.png")
 ```
 
-From these tables, we can see that the R^2's for these models are still relatively low but a simple linear regression of using percent vaccination or the location of each country have R^2's of around 0.35 which is an improvement from the linear regression models from using logit transformation of difference in proportions as the outcome.
+Looking at Figure 2.4, this side-by-side boxplot shows that Africa is significantly different to all of the other continents.
+Africa is historically known is a less developed continent.
+This could be result in lower supplies to Africa compared to other places in the world.
+Within the other continents, Europe seems to have th highest median vaccinated proportions.
+
+```{r figure 2.4,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/loc_vs_vacc.png")
+```
+
+Figure 2.5 is showing the logit transformation of the proportion of the population vaccinated.
+This is due to the skewness of the non-transformed vaccinatio numbers.
+We see that this is still skewed but its skewed left where most of the data is around that 0 value.
+
+```{r figure 2.5,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/log_vacc.png")
+```
+
+#### Main Analysis
+
+As specified within the methods section, the first part of my analysis will be comparing the RMSE of my full model to the null model.
+However, before I do this I will run a correlation matrix of the numberic variables to see if some variables are highly correlated.
+
+In Table 2.2 we can see that the two variable, which I suspected, that are highly correlated are health care funds and GDP per capita where value is the health care fund variable.
+Since those two variables are highly correlated and the health care funds variable only includes 48 observations, I will remove health care funds out of the full model.
+
+Table 2.2
+```{r table 2.2,  echo=FALSE}
+tab22=readRDS("../../results/cor_tab.rds")
+knitr::kable(tab22)
+```
+
+Now that I have my full model solidified, Table 2.3 shows the summary table and some other statistics of this full model.
+We can see that our adjusted R-sqaured value is around 0.64 and most of our factors are significant except the proportion of this who died and the COVID tests per person.
+
+Table 2.3
+```{r table 2.3,  echo=FALSE}
+tab23=readRDS("../../results/Summ_stat_mult_reg.rds")
+knitr::kable(tab23)
+
+tab232=readRDS("../../results/Multiple_Reg_tab.rds")
+knitr::kable(tab232)
+
+```
+
+**Simple Comparison of RMSE**
+
+To start comparing the full model to the null model in terms of RMSE, we need to create training and test data sets.
+I used a 0.75 split for the training set.
+I am working on this project using Tidyverse and Tidymodels packages which uses the functions of recipes and workflows.
+The recipe I am specifying is the creation of the model which in this case, I need to make two.
+One of the recipes is for the full model and the other is for the null model.
+Since my outcome is numeric, we need to specify the type of model as well which is linear regression for our project.
+Once we have our model type and our recipe, we can combine them and put it in a workflow.
+This workflow will be able to combine the two `settings` we have and create a fit object which eventually creates the statistics of our models and our predictors.
+However, since our main objective is looking at the fit of our data, we want to look at the RMSE.
+
+We can fit our fit object with both our test and training sets to see how well it does for both data sets.
+A general consensus on RMSE is that models with a value close to 1 is good and those who go farther away in the positive direction have a worse fit.
+We want our test and train values to be similar as well as that shows the model does not overfit the training data and results in a lower fit on the test data.
+
+Table 2.4 shows the two values for the full model where the top is the training value and the bottom is the testing value.
+We can see that these two values are somewhat similar and possibly we want to a bit closer.
+However, these two values are relatively low which is a good sign.
+
+Comparing Table 2.4 to Table 2.5, which shows the null models results, we can see our full model does a lot better than the null model.
+The RMSE is a lot worse for the null model.
+This is a good sign for the data as we now know that these predictors do have some type of association since it is way better than the null model.
+We can now proceed to creating our LASSO model and tuning this full model to find the best performing model.
+
+Table 2.4
+```{r table 2.4,  echo=FALSE}
+tab24=readRDS("../../results/Logit_vacc_rmse_tab.rds")
+knitr::kable(tab24)
+```
+
+Table 2.5
+```{r table 2.5,  echo=FALSE}
+tab25=readRDS("../../results/Logit_vacc_rmse_null_tab.rds")
+knitr::kable(tab25)
+```
+
+**LASSO Model**
+
+For the LASSO model, there is some preparation we need to run this.
+We will use cross validation to create the multiple models included in the LASSO model.
+This cross validation will be a 5-fold stratified by the outcome.
+We then, like previously, create a workflow using a specific model type and recipe.
+The recipe will be the full model recipe but the model type will be different.
+The LASSO model uses a specific model that has a tuning factor into a linear regression type.
+Once we have our workflow, we need to make a grid as the LASSO model created multiple models based on the 5-fold cross validation and other factors.
+From the workflow, we can apply a tuning grid to it.
+This will produce the variety of models.
+To find the best model, we an apply a function called show_best and specify RMSE to find the best model in terms of RMSE.
+After putting this model into a workflow and creating a fit object, we get the statistics of the best model.
+Figure 2.6 shows the tuning process in a visual.
+The best model is represented by the lowest RMSE within the dip.
+
+```{r figure 2.6,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/lasso_tune_process.png")
+```
+
+Figure 2.7 shows the predicted values to the observed values.
+Most of the points follow the trend and there seems to show no pattern.
+We can see this more within Figure 2.8 which is the residual plot.
+Within Figure 2.8, we could see a slight pattern at the begging points where they are grouped up more.
+
+```{r figure 2.7,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/Pred_vs_obs_value_lasso_train.png")
+```
+
+```{r figure 2.8,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/Resid_lasso_train.png")
+```
+
+The last part of the LASSO modeling analysis is fitting the model to the test data to see how well they both look together.
+To look at the test data results, we need to run a last fit function which just applies the test data to the best model found above.
+Figure 2.9 and Figure 2.10 show the predicted values versus observed and its residual plot for the test data.
+Just like the training test, most of the predicted values do have a similar trend and it close to the observed value.
+The residual plot shows that concerning line of observations for the lower values.
+This residual plot show more variability in the lower values more.
+Table 2.6 shows the RMSE values created from this LASSO model.
+We can see unfortunately these are overall higher and farther apart.
+It is still better than the null model.
+
+```{r figure 2.9,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/Pred_vs_obs_value_lasso_test.png")
+```
+
+```{r figure 2.10,  fig.cap='Analysis figure.', echo=FALSE}
+knitr::include_graphics("../../results/Resid_lasso_test.png")
+```
+
+Table 2.6
+```{r table 2.6,  echo=FALSE}
+tab26=readRDS("../../results/LASSO_rmse_train_tab.rds")
+knitr::kable(tab26)
+
+tab262=readRDS("../../results/LASSO_rmse_test_tab.rds")
+knitr::kable(tab262)
+
+```
 

### 2021-11-12 products/manuscript/Supplemental Material.Rmd
MADA Project part 4

---
title: "MADA Project: Supplemental Material"
subtitle: ""
author: Dawson Dobash
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

A table summarizing the data.

```{r summarytable,  echo=FALSE}
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable, caption = 'Data summary table.')
```

A table of the top 10 countries with the highest percentage of cases by population. We can see that there are a variety of locations and that Montenegro and Czech Republic are the highest 2 countries with the most cases by population.

```{r table1,  echo=FALSE}
tab1=readRDS("../../results/table1.rds")
knitr::kable(tab1)
```

A table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate. For the bottom 5, there is a mixture of locations. We can also see that the recovery rate and death rate for the bottom 5 teams do not add up to 100 or close to it. This means there are a lot of active cases.

```{r table2,  echo=FALSE}
tab2=readRDS("../../results/table2.rds")
knitr::kable(tab2)
```

A table  of the top 5 and bottom 5 countries for death rate. We can see that Vanuatu only has a total of 4 cases but Yemen also has a significantly large rate then the next top country, Mexico. We can also see that Asia dominates the lowest death rates.

```{r table3,  echo=FALSE}
tab3=readRDS("../../results/table3.rds")
knitr::kable(tab3)
```

A scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.

```{r figure1,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure1.png")
```

A scatterplot figure of total cases by population filtered down to countries with less than 40 million population. We can see that the total cases are spread out.

```{r figure2,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure2.png")
```

A scatterplot figure of total recovered patients by total cases. We can see just like in figure 1 that the United States, India, and Brazil have a overwhelmingly more cases than other countries. Therefore in figure 4, we will explore the majority of the countries.

```{r figure3,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure3.png")
```

A scatterplot figure of the same plot as before just filtering countries with less than 20 million cases. We can see a linear relationship between the recovered to total cases. It looks like a slope of 1.

```{r figure4,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure4.png")
```

A scatterplot figure of percent of people fully vaccinated by health care funding. We can see that it is a linear relationship with a positive correlation. We can also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.

```{r figure6,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure6.png")
```

A histogram figure of the percentage of cases per population. We can see that this data is highly skewed to the right.

```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
knitr::include_graphics("../../results/figure8.png")
```

We can see with looking at figure 5 that the percentage of cases are roughly bimodal or even right skewed. For now, I will proceed with doing some simple linear regression. Below shows model statistics for each model and a visual showing the relationship between the two variables.

Figure 10 is showing statistics for the predictor percentage of cases.
```{r vaccine pct,  echo=FALSE}
knitr::include_graphics("../../results/figure10.png")

tab9=readRDS("../../results/tablepctcases2.rds")
knitr::kable(tab9)
```

Figure 11 is showing statistics for the predictor health care funds.
```{r health care,  echo=FALSE}
knitr::include_graphics("../../results/figure11.png")

tab10=readRDS("../../results/tablehc2.rds")
knitr::kable(tab10)
```

Figure 12 is showing statistics for the predictor number of tests per person.
```{r Tests,  echo=FALSE}
knitr::include_graphics("../../results/figure12.png")

tab11=readRDS("../../results/tabletests2.rds")
knitr::kable(tab11)
```

Figure 13 is showing statistics for the predictor location which specifies the continent for each country.
```{r Location,  echo=FALSE}
knitr::include_graphics("../../results/figure13.png")

tab12=readRDS("../../results/tableloc2.rds")
knitr::kable(tab12)
```

### 2021-11-12 code/analysis_code/analysisscript.R

@@ -32,14 +32,14 @@ summarytable = summary(mydata)
 
 #For this scatter plot, I plotted each countries population compare to their total cases of COVID-19.
 #You can see that there are three countries that either has more population or total cases than others
-fig1 = mydata %>% ggplot(aes(x=Population, y=`Total Cases`, label=Country)) + geom_point() + 
+case_vs_pop = mydata %>% ggplot(aes(x=Population, y=`Total Cases`, label=Country)) + geom_point() + 
   geom_text(aes(label=ifelse(`Total Cases`>20000000, as.character(Country),'')), hjust=.3,vjust=1) + 
-            ggtitle("Figure 1: Scatterplot of Total Cases by Population for each Country \n Labeling some outliers")
+            ggtitle("Scatterplot of Total Cases by Population for each Country \n Labeling some outliers")
 
 #This scatter plot is in response to looking at only countries that has less than 40 million people. This is to
 #show the distribution of the other countries.
-fig2 = mydata %>% filter(Population <= 40000000) %>% ggplot(aes(x=Population, y=`Total Cases`)) + geom_point() + 
-  ggtitle("Figure 2: Scatterplot of Total Cases by Population for each Country \n under the Population of 40 million")
+case_vs_pop2 = mydata %>% filter(Population <= 40000000) %>% ggplot(aes(x=Population, y=`Total Cases`)) + geom_point() + 
+  ggtitle("Scatterplot of Total Cases by Population for each Country \n under the Population of 40 million")
 
 #I wanted to see the top 10 Countries with the most cases proportionally to their population. I included the 
 #continent that are on just to see if there was any connection. We can see that there is a good mixture of countries 
@@ -51,16 +51,16 @@ tab1 = mydata %>% mutate(totcase_percent = (`Total Cases`/Population)*100) %>% a
 #that the same three countries have much larger numbers than all of the other ones. Just by the look of this
 #It looks to have a simple linear relationship. I wanted to see if there were any countries that could be doing the 
 #recovery/health care process.
-fig3 = mydata %>% ggplot(aes(y=`Total Recovered`, x=`Total Cases`, label = Country)) + geom_point() + 
[…]

### 2021-11-12 code/analysis_code/main_analysis.R

@@ -1,0 +1,394 @@
+##Main Analysis
+
+#load needed packages. make sure they are installed.
+library(ggplot2) #for plotting
+library(broom) #for cleaning up output from lm()
+library(here) #for data loading/saving
+library(tidyverse)
+library(AER)
+library(tidymodels)
+library(dotwhisker)
+
+
+
+
+#path to data
+#note the use of the here() package and not absolute paths
+data_location <- here::here("data","processed_data","processeddata.rds")
+
+#load data. 
+mydata <- readRDS(data_location)
+
+#Setting seed
+set.seed(345)
+
+######### Analysis Part 2- percentage of vaccinated people within a country ###############
+
+## Correlation Matrix of all predictors ##
+
+model_num_data <- mydata %>% select(prop_vacc, pct_cases, test_per_person, Value, gdp_per_capita)
+cor_tab = cor(model_num_data, use = "complete.obs")
+cor_file = here("results", "cor_tab.rds")
+saveRDS(cor_tab, file = cor_file)
+
+
+## Since health care funds (Value) and gdp have a correlation of 0.87 and there are a lot of missing data within health care funds,
+## I will remove health care funds and replace it with gdp for my full model
+
+
+
+
+
+####################################################################
+######### Visualization of main outcome to predictors ##############
+####################################################################
+
+#Visualization of the simple linear regression models
+
+#This is a table showing the top 5 and bottom 5 percentage of vaccinated people 
+vacc_tab = mydata %>% select(Country, location, `% of population vaccinated`) %>% 
+  arrange(desc(`% of population vaccinated`)) %>% {rbind(head(., 5), tail(., 5))}
+vacc_file = here("results", "vacc_tab.rds")
+saveRDS(vacc_tab, file = vacc_file)
+
+#This is a histogram showing the distribution of those who are vaccinated. We can see that it is somewhat uniform
+vacc_hist = mydata %>% ggplot(aes(x = `% of population vaccinated`)) + geom_histogram(binwidth = 5) + 
+  ggtitle("Figure 2.1: Histogram of Percent of People Vaccinated \n (by Country)")
[…]

### 2021-11-12 code/processing_code/processingscript.R

@@ -23,6 +23,8 @@ data_location5 <- here::here("data","raw_data","Covid Data - Europe.csv")
 data_location6 <- here::here("data","raw_data","Covid Data - North America.csv")
 data_location7 <- here::here("data","raw_data","Covid Data - Oceania.csv")
 data_location8 <- here::here("data","raw_data","Covid Data - South America.csv")
+data_location9 <- here::here("data","raw_data","GDP per capita.csv")
+data_location10 <- here::here("data","raw_data","Country_Govt.csv")
 
 #load data from the data locations specified for each data set 
 
@@ -34,6 +36,8 @@ europe <- readr::read_csv(data_location5)
 n_america <- readr::read_csv(data_location6)
 oceania <- readr::read_csv(data_location7)
 s_america <- readr::read_csv(data_location8)
+gdp <- readr::read_csv(data_location9)
+govt <- readr::read_csv(data_location10)
 
 
 #take a look at the data.I only included looking at the africa data set because the other continent data sets have the
@@ -92,6 +96,22 @@ clean_funds2$Country = countrycode(clean_funds2$LOCATION, "iso3c", "country.name
 clean_funds2$Country = ifelse(clean_funds2$Country == "Czechia", "Czech Republic", 
                               ifelse(clean_funds2$Country == "United Kingdom", "UK", clean_funds2$Country))
 
+gdp <- gdp %>% mutate(ifelse(Country == "United Kingdom", "UK", 
+                             ifelse(Country == "Bahamas, The", "Bahamas", 
+                                    ifelse(Country == "Brunei Darussalam", "Brunei", 
+                                           ifelse(Country == "Cabo Verde", "Cape Verde", 
+                                                  ifelse(Country == "Congo, Dem. Rep.", "Congo", 
+                                                         ifelse(Country == "Egypt, Arab Rep.", "Egypt", 
+                                                                ifelse(Country == "Gambia, The", "Gambia", 
+                                                                       ifelse(Country == "Hong Kong SAR, China", "Hong Kong", 
+                                                                              ifelse(Country == "Iran, Islamic Rep.", "Iran", 
+                                                                                     ifelse(Country == "Cote d'Ivoire", "Ivory Coast", 
+                                                                                            ifelse(Country == "Congo, Rep.", "Republic of the Congo", 
+                                                                                                   ifelse(Country == "Russian Federation", "Russia", 
[…]

### 2021-11-12 products/manuscript/Manuscript.Rmd
MADA project part 4-added abstract

@@
 
-# Summary/Abstract
-_Write a summary of your project._
+# Abstract
+COVID-19 has been the talk of the world for the past two years, and it is still prevalent.
+The affects of COVID-19 not only affected individual people, but it also affected businesses, health protocols, and the economy of governments.
+It has affected all countries not only specific to a type of country which is the scary part.
+It is truly indeed a pandemic.
+
+I am interested in looking at the difference between the proportion of those who recovered to those who have died for each country in the world.
+I am also looking at how the proportion of people vaccinated in each country is affected by multiple factors.
+The predictors I will be considering are: percent of cases within a population, continent the country is on, number of tests per person, gdp per capita, proportion of those who died, government type, and health care funds.
+
+By looking at some simple linear regressions and other exploratory analysis, I determined due to how high the recovery rate is within each country overall and no association based on R-squared values, the difference in proportions did not change based on multiple factors.
+
+While looking at vaccination proportions, I decided to do some analyses using training/test data sets and applying them to find the RMSE of the full model and a null model.
+When finding the RMSE, I found that the full model does a better job in fitting the data compared to the null model.
+This is a good sign to share that some combination of predictors have an association to the proportion of vaccinated people for each country.
+
+The next analysis was creating a machine learning modeling technique called the LASSO model.
+This type of modeling will create multiple models from a five-fold cross-validation and a tuning grid which will try to find the best performing model.
+From the LASSO we find that model 25 give the best model.
+
+**Need to find a way to extract this model**
 

### 2021-11-28 products/manuscript/Manuscript.Rmd
MADA project part 5 (no references yet)

@@
 ---
-title: "MADA Project Manuscript"
+title: "MADA Project: COVID-19 Vaccinations and recovery"
 subtitle: ""
@@
 
-https://data.oecd.org/healthres/health-spending.htm
+[https://data.oecd.org/healthres/health-spending.htm](https://data.oecd.org/healthres/health-spending.htm)
 
-https://www.kaggle.com/iamsouravbanerjee/covid19-dataset-world-and-continent-wise?select=Covid+Data+-+World.csv
+[https://www.kaggle.com/iamsouravbanerjee/covid19-dataset-world-and-continent-wise?select=Covid+Data+-+World.csv](https://www.kaggle.com/iamsouravbanerjee/covid19-dataset-world-and-continent-wise?select=Covid+Data+-+World.csv)
 
-https://www.kaggle.com/anandhuh/latest-worldwide-vaccine-data
+[https://www.kaggle.com/anandhuh/latest-worldwide-vaccine-data](https://www.kaggle.com/anandhuh/latest-worldwide-vaccine-data)
 
-https://data.worldbank.org/indicator/NY.GDP.PCAP.CD
+[https://data.worldbank.org/indicator/NY.GDP.PCAP.CD](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD)
 
-https://en.wikipedia.org/wiki/List_of_countries_by_system_of_government
+[https://en.wikipedia.org/wiki/List_of_countries_by_system_of_government](https://en.wikipedia.org/wiki/List_of_countries_by_system_of_government)
 
@@
 
-## Full analysis
-
-### Difference in proportions of recovery and deaths
+## Question 1: Difference in proportions of recovery and deaths
 
@@
 We can see that the bottom five death rates are from Asia but also the highest, Yemen, is also within the Asia continent as well.
+<br /> 
+<br />
 
-Table 1.1
+Table 1.1: Top 5 and bottom 5 death rates by country
 ```{r table1,  echo=FALSE}
@@
 
-```{r figure 1.1,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 1.1, echo=FALSE}
 knitr::include_graphics("../../results/prop_diff_recov_died.png")
@@
 
-```{r figure 1.2,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 1.2, echo=FALSE}
 knitr::include_graphics("../../results/log_prop_diff_hist.png")
@@
 
-Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of each of the models within Tables 1.2-1.5.
+Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of one of the models within Tables 1.2.
 
@@
 
-Table 1.2
+Table 1.2: Summary statistics for linear regression (logit transformation of prop diff of recovery and deaths vs. vaccine)
 ```{r table 1.2,  echo=FALSE}
@@
 
-Table 1.3 is showing statistics for the predictor health care funds.
-
-Table 1.3
-```{r table 1.3,  echo=FALSE}
-tab3=readRDS("../../results/tablehc1.rds")
-knitr::kable(tab3)
-```
-
-Table 1.4 is showing statistics for the predictor number of tests per person.
-
-Table 1.4
-```{r table 1.4,  echo=FALSE}
-tab4=readRDS("../../results/tabletests1.rds")
-knitr::kable(tab4)
-```
-
-Table 1.5 is showing statistics for the predictor location which specifies the continent for each country.
-
-Table 1.5
-```{r table 1.5,  echo=FALSE}
-tab5=readRDS("../../results/tableloc1.rds")
-knitr::kable(tab5)
-```
-
-Looking at all of these tables, we can see that the R^2 values are very low.
+Looking at Table 1.2, we can see that the R^2 values are very low.
+This is similar to all of the other simple linear regression models.
+The statistics for the other three models are within the supplemental material.
 Therefore none of these predictors shows any association with the logit transformation outcome.
@@
 
-### Percent of people vaccinated for each country
+## Question 2: Percent of people vaccinated for each country
 
-#### Methods
+### Methods
 
@@
 
-#### Exploratory Analysis
+### Exploratory Analysis
 
@@
 
-Table 2.1
+Table 2.1: Top 5 and bottom 5 vaccination percentages by country
 ```{r table 2.1,  echo=FALSE}
@@
 
-```{r figure 2.1,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.1, echo=FALSE}
 knitr::include_graphics("../../results/vacc_hist.png")
@@
 
-Figure 2.2 shows that there is no trend between vaccinations and cases until you get out to over 10% of the population had cases of COVID-19.
-The vaccination numbers range a lot when there is a low perecentage of cases compared to the countries population.
+When looking at Figure 2.2, we can see that those countries who have large values of GDP per capita, their population is more vaccinated.
+The vaccination number varies when the GDP per capita is relatively low.
 
-```{r figure 2.2,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/cases_vs_vacc.png")
-```
-
-When looking at Figure 2.3, we can see that those countries who have large values of GDP per capita, their population is more vaccinated.
-The vaccination number, just like in Figure 2.2, varies when the GDP per capita is relatively low.
-
-```{r figure 2.3,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.2, echo=FALSE}
 knitr::include_graphics("../../results/gdp_vs_vacc.png")
@@
 
-Looking at Figure 2.4, this side-by-side boxplot shows that Africa is significantly different to all of the other continents.
+Looking at Figure 2.3, this side-by-side boxplot shows that Africa is significantly different to all of the other continents.
 Africa is historically known is a less developed continent.
@@
 
-```{r figure 2.4,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.3, echo=FALSE}
 knitr::include_graphics("../../results/loc_vs_vacc.png")
@@
 
-Figure 2.5 is showing the logit transformation of the proportion of the population vaccinated.
+Figure 2.4 is showing the logit transformation of the proportion of the population vaccinated.
 This is due to the skewness of the non-transformed vaccinatio numbers.
@@
 
-```{r figure 2.5,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.4, echo=FALSE}
 knitr::include_graphics("../../results/log_vacc.png")
@@
 
-#### Main Analysis
+### Main Analysis
 
 As specified within the methods section, the first part of my analysis will be comparing the RMSE of my full model to the null model.
-However, before I do this I will run a correlation matrix of the numberic variables to see if some variables are highly correlated.
+However, before I do this I will run a correlation matrix of the numeric variables to see if some variables are highly correlated.
 
-In Table 2.2 we can see that the two variable, which I suspected, that are highly correlated are health care funds and GDP per capita where value is the health care fund variable.
+In Table 2.2 we can see that the two variable, which I suspected, that are highly correlated are health care funds (Value) and GDP per capita.
 Since those two variables are highly correlated and the health care funds variable only includes 48 observations, I will remove health care funds out of the full model.
 
-Table 2.2
+Table 2.2: Correlation martrix of predictors and outcome
 ```{r table 2.2,  echo=FALSE}
@@
 
-Table 2.3
+Table 2.3: Summary statistics for full model
 ```{r table 2.3,  echo=FALSE}
@@
 
-Table 2.4
+Table 2.4: RMSE values for full model (training and test data, respectively)
 ```{r table 2.4,  echo=FALSE}
@@
 
-Table 2.5
+Table 2.5: RMSE values for null model (training and test data, respectively)
 ```{r table 2.5,  echo=FALSE}
@@
 After putting this model into a workflow and creating a fit object, we get the statistics of the best model.
-Figure 2.6 shows the tuning process in a visual.
-The best model is represented by the lowest RMSE within the dip.
 
-```{r figure 2.6,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/lasso_tune_process.png")
-```
+Figure 2.5 shows the predicted values to the observed values.
+Most of the points follow the trend and there seems to show no pattern.
+We can see this more within Figure 2.6 which is the residual plot.
+Within Figure 2.6, we could see a slight pattern at the begging points where they are grouped up more.
 
-Figure 2.7 shows the predicted values to the observed values.
-Most of the points follow the trend and there seems to show no pattern.
-We can see this more within Figure 2.8 which is the residual plot.
-Within Figure 2.8, we could see a slight pattern at the begging points where they are grouped up more.
-
-```{r figure 2.7,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.5, echo=FALSE}
 knitr::include_graphics("../../results/Pred_vs_obs_value_lasso_train.png")
@@
 
-```{r figure 2.8,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.6, echo=FALSE}
 knitr::include_graphics("../../results/Resid_lasso_train.png")
@@
 To look at the test data results, we need to run a last fit function which just applies the test data to the best model found above.
-Figure 2.9 and Figure 2.10 show the predicted values versus observed and its residual plot for the test data.
+Figure 2.7 and Figure 2.8 show the predicted values versus observed and its residual plot for the test data.
 Just like the training test, most of the predicted values do have a similar trend and it close to the observed value.
@@
 Table 2.6 shows the RMSE values created from this LASSO model.
-We can see unfortunately these are overall higher and farther apart.
-It is still better than the null model.
+We can see unfortunately these are overall higher compared to the full model numbers.
+It is still better than the null model however.
 
-```{r figure 2.9,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.7, echo=FALSE}
 knitr::include_graphics("../../results/Pred_vs_obs_value_lasso_test.png")
@@
 
-```{r figure 2.10,  fig.cap='Analysis figure.', echo=FALSE}
+```{r figure 2.8, echo=FALSE}
 knitr::include_graphics("../../results/Resid_lasso_test.png")
@@
 
-Table 2.6
+Table 2.6: Comparing RMSE for training and test split data for LASSO model, respectively 
 ```{r table 2.6,  echo=FALSE}
@@
 
+This is the product from the LASSO modeling.
+This is the model produced by LASSO.
+In Table 2.7, we can see there are 4 predictors and 2 were removed.
+The predictors that were removed was the proportion of those who died in each country and test per person.
+However these estimates are based on the logit transformation of our outcome.
+The last column is the transformation to display the transformed of each predictor applying the exp() function to the estimates.
+
+From the estimates, we can see although there was 2 variables that suppressed to a value of 0 based on the LASSO model, the predictor gdp per capita have roughly a value of 0 which translates to a odds ratio close to 1.
+Those who include 1 or close to 1 results in a non-significant variable in predicting the outcome.
+We can see the continent that the country is in also has an affect on proportion vaccinated where the continent Africa is the reference group.
+The continents that increases the proportion of being vaccinated the most are Asia and South America.
+We can also see that in terms of government type where a republic government is the reference group, those countries that are a monarchy increase the proportion of having more vaccinations.
+In opposite affect, a provisional government country lowers the proportion of getting people vaccinated.
+
+Table 2.7: Summary statistics of LASSO model and predictors with their respected transformed values 
+```{r table 2.7,  echo=FALSE}
+tab27=readRDS("../../results/transform_LASSO.rds")
+knitr::kable(tab27)
+```
+
 # Discussion
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+The analysis was split up into two different analyses specified by the two questions.
+The first analysis was looking at the difference of proportions between recovery and death rates.
+With some exploratory analysis, no association according to the simple linear regression between multiple predictors and the outcome.
+Since within some exploratory analysis and very small r-square values, I decided to stop that analysis on the first question and go to the next question.
+
+<\ br>
+
+The second question pertains to vaccination proportions.
+Using the logit transformation and applying the LASSO modeling, out of the 6 predictors within the full model, only 4 remained within the equation.
+These include percent of cases for a country, gdp per capita, the continent the country resides, and the type of government.
+Out of these, those who have the most impact are location and government type, and gdp per capita has roughly an estimate of 0 which probably has no impact on the proportion of people vaccinated.
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+Some strengths to this analysis was that since this topic of COVID-19 is in todays news which means much of the countries information is available in public use.
+Along with COIVD-19 information, most information of countries are of public use therefore lots of information  was easy to find.
+
+Some limitations for this analysis was that I was using cumulative data which means as more days pass, this data will become old analyses.
+Another limitation was although I got the information I wanted, I had to use multiple sources to find them which results in combining data sets and ultimately errors.
 
 ## Conclusions
-_What are the main take-home messages?_
+We find that due to high high recovery right is for most countries compared to death rates, and no predictors have a large association, according to the simple linear regression r-squared values, the difference in proportion between recovery and death rates is not affected by most predictors.
+
+We see that the proportion of people vaccinated is associated with the amount of COVID-19 cases, which continent the country is in, and type of government.
+Given all other variable constant, Asia and South America gives the highest proportional increase compared to Africa, and monarchy governments increase vaccinations compared to republic government countries.
 

### 2021-11-28 products/manuscript/Supplemental Material.Rmd
MADA project part 5 (no references yet)

@@
 ---
+This document displays some supplemental material to the analysis.
 
-A table summarizing the data.
+**Supplemental material for question 1**
 
-```{r summarytable,  echo=FALSE}
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable, caption = 'Data summary table.')
-```
+This table is showing statistics for the simple linear regression between the predictor health care funds with the outcome being the logit transformation of the proportional difference between recovery and death rates.
+This is supplemented for table 1.2.
 
-A table of the top 10 countries with the highest percentage of cases by population.
-We can see that there are a variety of locations and that Montenegro and Czech Republic are the highest 2 countries with the most cases by population.
+This helps describe that no predictors are good in predicting the outcome with a low R-squared.
 
-```{r table1,  echo=FALSE}
-tab1=readRDS("../../results/table1.rds")
-knitr::kable(tab1)
-```
-
-A table of the top 5 and bottom 5 recovery rates also sharing the total cases numbers.It looks that Asia dominates the top 5 countries for recovery rate.
-For the bottom 5, there is a mixture of locations.
-We can also see that the recovery rate and death rate for the bottom 5 teams do not add up to 100 or close to it.
-This means there are a lot of active cases.
-
-```{r table2,  echo=FALSE}
-tab2=readRDS("../../results/table2.rds")
-knitr::kable(tab2)
-```
-
-A table  of the top 5 and bottom 5 countries for death rate.
-We can see that Vanuatu only has a total of 4 cases but Yemen also has a significantly large rate then the next top country, Mexico.
-We can also see that Asia dominates the lowest death rates.
-
-```{r table3,  echo=FALSE}
-tab3=readRDS("../../results/table3.rds")
+```{r table regr for hc,  echo=FALSE}
+tab3=readRDS("../../results/supp_mat/tablehc1.rds")
 knitr::kable(tab3)
@@
 
-A scatterplot figure  of total cases by population. United States, India, and Brazil are some outliers of the countries.
+This table is showing statistics for the simple linear regression between the predictor number of tests performed with the outcome being the logit transformation of the proportional difference between recovery and death rates.
+This is supplemented for table 1.2.
 
-```{r figure1,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure1.png")
+This helps describe that no predictors are good in predicting the outcome with a low R-squared.
+
+```{r table regr for tests,  echo=FALSE}
+tab4=readRDS("../../results/supp_mat/tabletests1.rds")
+knitr::kable(tab4)
 ```
 
-A scatterplot figure of total cases by population filtered down to countries with less than 40 million population. We can see that the total cases are spread out.
+This table is showing statistics for the simple linear regression between the predictor location, which is the continent each country is associated with, with the outcome being the logit transformation of the proportional difference between recovery and death rates.
+This is supplemented for table 1.2.
 
-```{r figure2,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure2.png")
+This helps describe that no predictors are good in predicting the outcome with a low R-squared.
+
+```{r table regr for loc,  echo=FALSE}
+tab5=readRDS("../../results/supp_mat/tableloc1.rds")
+knitr::kable(tab5)
 ```
 
-A scatterplot figure of total recovered patients by total cases.
-We can see just like in figure 1 that the United States, India, and Brazil have a overwhelmingly more cases than other countries.
-Therefore in figure 4, we will explore the majority of the countries.
+**Supplemental material for question 2**
 
-```{r figure3,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure3.png")
-```
+This table is showing statistics for the simple linear regression between the predictor percentage of cases with the outcome being the logit transformation of the proportion of people vaccinated.
 
-A scatterplot figure of the same plot as before just filtering countries with less than 20 million cases.
-We can see a linear relationship between the recovered to total cases.
-It looks like a slope of 1.
-
-```{r figure4,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure4.png")
-```
-
-A scatterplot figure of percent of people fully vaccinated by health care funding.
-We can see that it is a linear relationship with a positive correlation.
-We can also see that the United States is out of the norm where they spend a lot more on health care funding compared to the percent of people fully vaccinated.
-
-```{r figure6,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure6.png")
-```
-
-A histogram figure of the percentage of cases per population. We can see that this data is highly skewed to the right.
-
-```{r figure8,  fig.cap='Analysis figure.', echo=FALSE}
-knitr::include_graphics("../../results/figure8.png")
-```
-
-We can see with looking at figure 5 that the percentage of cases are roughly bimodal or even right skewed.
-For now, I will proceed with doing some simple linear regression.
-Below shows model statistics for each model and a visual showing the relationship between the two variables.
-
-Figure 10 is showing statistics for the predictor percentage of cases.
 ```{r vaccine pct,  echo=FALSE}
-knitr::include_graphics("../../results/figure10.png")
-
-tab9=readRDS("../../results/tablepctcases2.rds")
+tab9=readRDS("../../results/supp_mat/tablepctcases2.rds")
 knitr::kable(tab9)
@@
 
-Figure 11 is showing statistics for the predictor health care funds.
+This table is showing statistics for the simple linear regression between the predictor GDP per capita with the outcome being the logit transformation of the proportion of people vaccinated.
+
 ```{r health care,  echo=FALSE}
-knitr::include_graphics("../../results/figure11.png")
-
-tab10=readRDS("../../results/tablehc2.rds")
+tab10=readRDS("../../results/supp_mat/tablegdp2.rds")
 knitr::kable(tab10)
@@
 
-Figure 12 is showing statistics for the predictor number of tests per person.
+This table is showing statistics for the simple linear regression between the predictor number of COVID-19 tests per person with the outcome being the logit transformation of the proportion of people vaccinated.
+
 ```{r Tests,  echo=FALSE}
-knitr::include_graphics("../../results/figure12.png")
-
-tab11=readRDS("../../results/tabletests2.rds")
+tab11=readRDS("../../results/supp_mat/tabletests2.rds")
 knitr::kable(tab11)
@@
 
-Figure 13 is showing statistics for the predictor location which specifies the continent for each country.
+This table is showing statistics for the simple linear regression between the predictor location which is the continent with the outcome being the logit transformation of the proportion of people vaccinated.
+
 ```{r Location,  echo=FALSE}
-knitr::include_graphics("../../results/figure13.png")
-
-tab12=readRDS("../../results/tableloc2.rds")
+tab12=readRDS("../../results/supp_mat/tableloc2.rds")
 knitr::kable(tab12)
 ```
+
+This figure shows that there is no trend between vaccinations and cases until you get out to over 10% of the population had cases of COVID-19.
+The vaccination numbers range a lot when there is a low perecentage of cases compared to the countries population.
+
+```{r case vs vacc, echo=FALSE}
+knitr::include_graphics("../../results/supp_mat/cases_vs_vacc.png")
+```
+
+This figure shows the tuning process in a visual. The best model is represented by the lowest RMSE within the dip.
+
+```{r tune process, echo=FALSE}
+knitr::include_graphics("../../results/supp_mat/lasso_tune_process.png")
+```

### 2021-11-28 code/analysis_code/analysisscript.R

@@ -88,7 +88,8 @@ vacc_vs_hc = mydata %>% ggplot(aes(x=Value, y=`% of population fully vaccinated`
 #This is a histogram of the difference in proportions of those who are recovered and who died with each denominator as
 #number of cases.
 prop_diff_recov_died = mydata %>% ggplot(aes(x = prop_diff_recov_vs_death )) + geom_histogram(binwidth = 0.05) + 
-  ggtitle("Figure 1.1: Histogram of Difference in proportions \n of those who recovered versus those who died \n (by Country)")
+  ggtitle("Figure 1.1: Histogram of Difference in proportions \n of those who recovered versus those who died \n (by Country)") + 
+  xlab("Proportion difference between recovery rates and death rates")
 
 
 #Looking at the distribution using a histogram
@@ -164,7 +165,8 @@ mydata$logit_prop_diff = logit(mydata$prop_diff_recov_vs_death)
 #This is a histogram of the logit difference in proportions of those who are recovered and who died with each denominator as
 #number of cases.
 log_prop_diff_hist = mydata %>% ggplot(aes(x = logit_prop_diff )) + geom_histogram(binwidth = 0.5) + 
-  ggtitle("Figure 1.2: Histogram of logit transformation of difference in proportions \n of those who recovered versus those who died \n (by Country)")
+  ggtitle("Figure 1.2: Histogram of logit transformation of difference \n in proportions of those who recovered versus those who died \n (by Country)") + 
+  xlab("Logit transformation of the proportional difference between recovory and death rates")
 log_prop_diff_hist
 
 #Saving figure * to a location for future references

### 2021-11-28 code/analysis_code/main_analysis.R

@@ -19,9 +19,6 @@ data_location <- here::here("data","processed_data","processeddata.rds")
 #load data. 
 mydata <- readRDS(data_location)
 
-#Setting seed
-set.seed(345)
-
 ######### Analysis Part 2- percentage of vaccinated people within a country ###############
 
 ## Correlation Matrix of all predictors ##
@@ -60,13 +57,13 @@ ggsave(filename = figvacc_file, plot= vacc_hist)
 
 #Scatterplots of these simple linear regressions
 case_scatterplot <- mydata %>% ggplot(aes(y=prop_vacc, x = pct_cases)) + geom_point() + 
-  ggtitle("Figure 2.2: Scatterplot of percentage of cases and \n proportion of people vaccinated \n (by Country)") 
+  ggtitle("Scatterplot of percentage of cases and \n proportion of people vaccinated \n (by Country)") 
 fig10_file = here("results","cases_vs_vacc.png")
 ggsave(filename = fig10_file, plot=case_scatterplot)
 
 gdp_per_cap_scatterplot <- mydata %>% ggplot(aes(x=gdp_per_capita, y = prop_vacc)) + geom_point() + 
-  ggtitle("Figure 2.3: Scatterplot of GPD per capita and proportion of people vaccinated \n (by Country)") +
-  xlab("GDP_per_capita")
+  ggtitle("Figure 2.2: Scatterplot of GPD per capita and proportion of \n people vaccinated (by Country)") +
+  xlab("GDP per capita") + ylab("Proportion vaccinated")
 fig11_file = here("results","gdp_vs_vacc.png")
 ggsave(filename = fig11_file, plot=gdp_per_cap_scatterplot)
 
@@ -80,8 +77,8 @@ ggsave(filename = fig12_file, plot=tests_scatterplot2)
 
 #Simple boxplot of proportion of population vaccinated by continent
 loc_boxplot <- mydata %>% ggplot(aes(x=location, y = prop_vacc)) + geom_boxplot()  + 
-  ggtitle("Figure 2.4: Boxplot of the proportion of people vaccinated by location \n (for each Country)") +
-  xlab("Continent")
+  ggtitle("Figure 2.3: Boxplot of the proportion of people vaccinated \n by location (for each Country)") +
+  xlab("Continent") + ylab("Proportion vaccinated")
 fig13_file = here("results","loc_vs_vacc.png")
 ggsave(filename = fig13_file, plot=loc_boxplot)
 
[…]

### 2021-11-28 code/processing_code/processingscript.R

@@ -11,6 +11,7 @@ library(dplyr) #for data processing
 library(here) #to set paths
 library(tidyverse) #for cleaning and 'tidying' the data
 library(countrycode) #changes country codes to full names
+library(gtools)
 
 #path to data
 #note the use of the here() package and not absolute paths

### 2021-11-29 products/manuscript/Manuscript.Rmd
MADA project part 5 (with references)

@@
 This type of modeling will create multiple models from a five-fold cross-validation and a tuning grid which will try to find the best performing model.
-From the LASSO we find that model 25 give the best model.
-
-**Need to find a way to extract this model**
+From the LASSO we find that model 27 give the best model, and the amount of cases, which continent the country resides in, and the type of government have impacts to the vaccination proportions.
+Compared to Africa, South America and Asia increases the proportion of vaccinations the most while an increase of cases means more vaccinations as well.
+A monarchy government compared to a republic government has more people vaccinated proportionally.
 
@@
 Other countries are still getting hit hard by this virus.
-In terms of vaccination output, there has been recent back and forth movements on people getting the vaccination.
+In terms of vaccination output, there has been back and forth movements on people getting the vaccination.
+A poll was close to when vaccinations were starting to disperse where only 51% of adults in the United States said they would get a vaccine [@choi21].
 Although in developed countries this happens but what about the less developed countries and regions such as Africa.
 While people in the United States are fighting about whether the vaccine should be given or not, the people in these other countries might have a limited supply.
+According to the United Nations in September, there is still an unequal amount of vaccinations going to developed and less developed countries.
+This could be due to the cost of each vaccination [@UN21].
+In a paper by Choi, as of January 2021, cases were fairly even between developed and less developed countries [@choi21].
+This just shows that although COVID-19 is affected everyone, the supply is not equal between all countries even though all countries need it.
 
 Looking at the difference in recovered patients and those who have died will be beneficial in determining if more care is needed in certain countries.
+According to  paper by Liang et al, the mortality rates were negatively impacted by the number of tests performed and also based on the countries economic status and government [@liang20].
 I also think the percentage of people vaccinated is also affected and can lead to important results that might show we need more either distribution or requirements of the vaccines.
@@
 
-<\ br>
-
 The second question pertains to vaccination proportions.
@@
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
 # References

### 2021-11-29 code/analysis_code/analysisscript.R
MADA project part 5 (updated read me files)

@@ -207,16 +207,16 @@ tests_stats = glance(tests_fit)
 loc_stats = glance(loc_fit)
 
 #Saving tables for later use
-tabhc1_file = here("results", "tablehc1.rds")
+tabhc1_file = here("results/supp_mat", "tablehc1.rds")
 saveRDS(hc_fund_stats, file = tabhc1_file)
 
 tabvacc1_file = here("results", "tablevacc1.rds")
 saveRDS(vacc_stats, file = tabvacc1_file)
 
-tabtests1_file = here("results", "tabletests1.rds")
+tabtests1_file = here("results/supp_mat", "tabletests1.rds")
 saveRDS(tests_stats, file = tabtests1_file)
 
-tabloc1_file = here("results", "tableloc1.rds")
+tabloc1_file = here("results/supp_mat", "tableloc1.rds")
 saveRDS(loc_stats, file = tabloc1_file)
 
 

### 2021-11-29 code/analysis_code/main_analysis.R

@@ -58,7 +58,7 @@ ggsave(filename = figvacc_file, plot= vacc_hist)
 #Scatterplots of these simple linear regressions
 case_scatterplot <- mydata %>% ggplot(aes(y=prop_vacc, x = pct_cases)) + geom_point() + 
   ggtitle("Scatterplot of percentage of cases and \n proportion of people vaccinated \n (by Country)") 
-fig10_file = here("results","cases_vs_vacc.png")
+fig10_file = here("results/supp_mat","cases_vs_vacc.png")
 ggsave(filename = fig10_file, plot=case_scatterplot)
 
 gdp_per_cap_scatterplot <- mydata %>% ggplot(aes(x=gdp_per_capita, y = prop_vacc)) + geom_point() + 
@@ -185,16 +185,16 @@ loc_stats3 = glance(loc_fit3)
 govt_stats = glance(govt_fit)
 
 #Saving tables for later use
-tabgdp2_file = here("results", "tablegdp2.rds")
+tabgdp2_file = here("results/supp_mat", "tablegdp2.rds")
 saveRDS(gdp_per_cap_stats3, file = tabgdp2_file)
 
-tabpctcases2_file = here("results", "tablepctcases2.rds")
+tabpctcases2_file = here("results/supp_mat", "tablepctcases2.rds")
 saveRDS(pctcases_stats3, file = tabpctcases2_file)
 
-tabtests2_file = here("results", "tabletests2.rds")
+tabtests2_file = here("results/supp_mat", "tabletests2.rds")
 saveRDS(tests_stats3, file = tabtests2_file)
 
-tabloc2_file = here("results", "tableloc2.rds")
+tabloc2_file = here("results/supp_mat", "tableloc2.rds")
 saveRDS(loc_stats3, file = tabloc2_file)
 
 tabgovt_file = here("results", "tablegovt.rds")
@@ -321,7 +321,7 @@ lr_res %>%
 
 #The autoplot function shows basically the process of the tuning process
 lasso_tune_plot <- lr_res %>% autoplot() + ggtitle("LASSO Tuning Process")
-lasso_tune_file = here("results", "lasso_tune_process.png")
+lasso_tune_file = here("results/supp_mat", "lasso_tune_process.png")
 ggsave(filename = lasso_tune_file, plot= lasso_tune_plot)
 
 #This is finding the best model with the lowest rmse

### 2021-12-02 Grunert_Peer_Review.Rmd
Ryan Grunert Peer Review
Completed the peer review for the project 12/1/21

# Overview

Title of project: MADA Project: COVID-19 Vaccinations and recovery

Name of project author(s): Dawson Dobash

Name of project reviewer: Ryan Grunert

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of the project is thoroughly described and the reason for investigating this issue is justified as well. Summaries of previous work are given, and are built upon to set the scene for this analysis. It is clear why we need to study COVID-19 vaccination and recovery in order to create some new insights into the spread and mortality rate of the disease.

Where I could see room for improvement, I suggest editing the introduction at the end in order to state the question clearly and the predictors/outcome you will be looking at. I gathered the information from the abstract initially, I think restating them at the end of the introduction will be beneficial for readers moving into the data description. My only other suggestion would be to edit for grammar throughout the introduction and abstract, it became difficult to figure out what points were being made with some sentences, seemed a little rushed when written. Other than that, great start to the MS.

### Summary assessment 
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions are clearly stated and looks like they can be answered using the data described in the previous section. Also states confounders that could influence results and answers to the questions. Only suggestion I have here is possibly expanding the "multiple predictors" in the second question by listing a few, but you mention the predictors in the earlier sections so I'm just being nitpicky. The goals for the analysis are clear.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Overall, the data is described well. All sources are provided, although no metadata file is available. Given, the majority of the variables are self-explanatory, but my suggestion would be adding a short metadata file that describes the proportions you created in the processing script. Other than that, your text in the MS regarding the data collection and description is great.

### Summary assessment

* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data is cleaned and processed well in the processing script. All of the steps are reproducible and well explained. No alternatives are discussed, but your method for wrangling was justified. You mentioned that you had to alter the csv file directly in order to alter the country names, but the use of the countrynames package was a good addition in order to fit the data together. The meaningful exploratory results are shown in the main manuscript file and are explained well. No qualms there. The supplementary materials are each explained as well. This section was done very well and I don't have much else to say to improve it.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis methods were appropriate for the data. All of the analysis is completely reproducible, so there are no problems there. Data preparation and model evaluation was done well. The majority of the code in the analysis scripts have comments that explain what each chunk is doing, so that is great also.

One suggestion I have is including what LASSO stands for, and why you are using it for the second question. I don't think you stated either, but I could have missed something. It would help add context for why you are using the model.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results are presented well and assist the reader in visualizing the data. I don't think they are publication level quality, but they get the job done. One suggestion I have is aligning the titles on the graphs in the middle, and giving a little more context with the tables. The tables present the needed information, although on some of them where there are multiple rows it becomes difficult to figure out which row pertains to which model, etc. Two ways to fix this could just be referencing the numbers for each model in the paragraphs, or altering the tables themselves to have an ID row (the RMSE tables for example). Overall though, well done on creating all the needed graphics.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The study findings are discussed, as well as some strengths and limitations, but not expanded on. Applying your findings to the population at large and focusing on any specific interesting findings you came across would be great additions to this section. The findings are stated and interpreted properly. My other suggestion will be expanded on in the further comments section below, but enough content is present in order to conclude the study successfully.

### Summary assessment

* minor parts wrong, missing or unclear

## Further comments

Overall, the project was finished and completely reproducible. The introduction, data collection, data processing and analysis, final manuscript, and supplementary materials are all present in an organized fashion. Code was concise and easy to run.

My main suggestion would be to go back through the manuscript, code scripts, and READMEs to edit the sentences and grammar. In multiple sections throughout the repository, it was difficult to understand what the content of some specific sentences were or the points that were being made. It looked like the majority of the writing was rushed, from my perspective. Going back to re-clarify your points and ideas would strengthen your overall analysis.

Other than that, great job! This project looked extensive with the multiple data sources and fitting them all together.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Overall, the project was well structured. All files are in well labeled folders and have reasonable names. All junk files not needed were removed, and I have a great idea of how things all fit together. The only suggestion I have for further improving the structure would be to create separate folders for Figures and Tables in the results folder to organize it a little more, but that's a little nitpicky. Structure of the code in each of the scripts is great as well.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is documented very well. I was able to understand each step of the entire analysis, each decision that was made, and almost every line of code. There was one section of code in the main_analysis script that didn't have comments explaining what it does, but I know enough about R to figure it out (multivariate regression). Adding in a couple comments there would eliminate that issue. There is enough information provided otherwise in the code or Rmd file for documentation of all the steps.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All the results are fully reproducible without any issues. READMEs provide enough information for what steps to take to reproduce the analysis, and no manual intervention is needed. I'm able to re-run the entire analysis without editing anything. Nothing to improve here.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study wasn't incredibly thorough, not many alternatives to processing or analyzing the data were discussed. A regression analysis was completed with univariate and multivariate models for the first question, but no other models or processing approaches were tried. My suggestion would be trying a machine learning model like a LASSO regression or a tree/forest just to see if you could get a new result (RMSE, etc.) from a different perspective, instead of just moving onto the next one right away. But it could be that you did, and didn't include it in the final repository. You could also try both bottom-up and top-down subset selection in the multivariate regression analysis, that's another option. Only the LASSO model was used for the second question, but it addressed it thoroughly enough to answer and conclude.

### Summary assessment

* decent level of thoroughness

## Further comments

Overall, very interesting topic and great project! Fully reproducible from beginning to end, and successfully analyzes data from multiple sources to answer the two questions. I think spending a little time addressing the above comments would be beneficial to the overall condition of the project.

### 2021-12-03 priyanka_peer_review.Rmd

--- 2021-12-02 Grunert_Peer_Review.Rmd
@@
+---
+title: Project Review Template 
+date: "`r file.mtime(knitr::current_input())`"
+#bibliography: ../media/references.bib
+output: 
+  html_document:
+    toc_depth: 3
+    number_sections: true
+---
+
 # Overview
 
-Title of project: MADA Project: COVID-19 Vaccinations and recovery
+Title of project:
 
@@
 
-Name of project reviewer: Ryan Grunert
+Name of project reviewer: Priyanka Gannavarapu
+
+# Instructions
+
+Write your comments and feedback below for each section/component of the project.
+The goal should be to help the author improve their project.
+Make comments as constructive and actionable as possible.
+You can provide both criticism and praise.
+
+For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.
+
+Make sure your final document compiles/knits into a readable, well-formatted html document.
+
+Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)
 
@@
 
-The context of the project is thoroughly described and the reason for investigating this issue is justified as well.
-Summaries of previous work are given, and are built upon to set the scene for this analysis.
-It is clear why we need to study COVID-19 vaccination and recovery in order to create some new insights into the spread and mortality rate of the disease.
+The project's context is clearly defined, as is the rationale for investigating this subject.
+The gaps from the previous studies are the best way to look for future studies.
+That’s the approach adopted here.
+It would be great to mention a clear objective.
+It seems a bit lost in the wording.
 
-Where I could see room for improvement, I suggest editing the introduction at the end in order to state the question clearly and the predictors/outcome you will be looking at.
-I gathered the information from the abstract initially, I think restating them at the end of the introduction will be beneficial for readers moving into the data description.
-My only other suggestion would be to edit for grammar throughout the introduction and abstract, it became difficult to figure out what points were being made with some sentences, seemed a little rushed when written.
-Other than that, great start to the MS.
-
-### Summary assessment 
+### Summary assessment (PICK ONE, DELETE THE OTHERS)
 * strong contextualization and motivation
@@
 
-The questions are clearly stated and looks like they can be answered using the data described in the previous section.
-Also states confounders that could influence results and answers to the questions.
-Only suggestion I have here is possibly expanding the "multiple predictors" in the second question by listing a few, but you mention the predictors in the earlier sections so I'm just being nitpicky.
-The goals for the analysis are clear.
+It's apparent what the author wants to know. It's evident from the questions that they're based on the data
 
 ### Summary assessment
-
 * question/hypotheses fully clear
@@
 
-Overall, the data is described well.
-All sources are provided, although no metadata file is available.
-Given, the majority of the variables are self-explanatory, but my suggestion would be adding a short metadata file that describes the proportions you created in the processing script.
-Other than that, your text in the MS regarding the data collection and description is great.
+The data is clearly described, as is the source.
 
 ### Summary assessment
-
-* source and overall structure of data somewhat explained
+* source and overall structure of data well explained
 
@@
 
-Data is cleaned and processed well in the processing script.
-All of the steps are reproducible and well explained.
-No alternatives are discussed, but your method for wrangling was justified.
-You mentioned that you had to alter the csv file directly in order to alter the country names, but the use of the countrynames package was a good addition in order to fit the data together.
-The meaningful exploratory results are shown in the main manuscript file and are explained well.
-No qualms there.
-The supplementary materials are each explained as well.
-This section was done very well and I don't have much else to say to improve it.
+Data cleaning was well done and all the steps were reproducible. All exploratory results are meaningful and presented well in the manuscript.
 
 ### Summary assessment
-
 * essentially no weaknesses in wrangling and exploratory component
@@
 
-The analysis methods were appropriate for the data.
-All of the analysis is completely reproducible, so there are no problems there.
-Data preparation and model evaluation was done well.
-The majority of the code in the analysis scripts have comments that explain what each chunk is doing, so that is great also.
-
-One suggestion I have is including what LASSO stands for, and why you are using it for the second question.
-I don't think you stated either, but I could have missed something.
-It would help add context for why you are using the model.
+The data analysis procedures were appropriate for the information presented. All the analysis is totally repeatable, including the results. it clearly shows what is being done and how it was done.
 
 ### Summary assessment
-
 * strong and reasonable analysis
@@
 
-The results are presented well and assist the reader in visualizing the data.
-I don't think they are publication level quality, but they get the job done.
-One suggestion I have is aligning the titles on the graphs in the middle, and giving a little more context with the tables.
-The tables present the needed information, although on some of them where there are multiple rows it becomes difficult to figure out which row pertains to which model, etc.
-Two ways to fix this could just be referencing the numbers for each model in the paragraphs, or altering the tables themselves to have an ID row (the RMSE tables for example).
-Overall though, well done on creating all the needed graphics.
+Results are presented well and has good flow to it. Serves the purpose.
 
 ### Summary assessment
-
-* results are presented ok, with room for improvement
+* results are very well presented
 
@@
 
-The study findings are discussed, as well as some strengths and limitations, but not expanded on.
-Applying your findings to the population at large and focusing on any specific interesting findings you came across would be great additions to this section.
-The findings are stated and interpreted properly.
-My other suggestion will be expanded on in the further comments section below, but enough content is present in order to conclude the study successfully.
+Discussion section can be worked on a little. Presenting the finding and discussing the previous literature would be a solution. Checking if the results are in line or different from previous studies
 
 ### Summary assessment
-
 * minor parts wrong, missing or unclear
@@
 
-Overall, the project was finished and completely reproducible.
-The introduction, data collection, data processing and analysis, final manuscript, and supplementary materials are all present in an organized fashion.
-Code was concise and easy to run.
-
-My main suggestion would be to go back through the manuscript, code scripts, and READMEs to edit the sentences and grammar.
-In multiple sections throughout the repository, it was difficult to understand what the content of some specific sentences were or the points that were being made.
-It looked like the majority of the writing was rushed, from my perspective.
-Going back to re-clarify your points and ideas would strengthen your overall analysis.
-
-Other than that, great job! This project looked extensive with the multiple data sources and fitting them all together.
+_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._
 
@@
 
-Overall, the project was well structured.
-All files are in well labeled folders and have reasonable names.
-All junk files not needed were removed, and I have a great idea of how things all fit together.
-The only suggestion I have for further improving the structure would be to create separate folders for Figures and Tables in the results folder to organize it a little more, but that's a little nitpicky.
-Structure of the code in each of the scripts is great as well.
+The project was completed and totally repeatable in its entirety.
+The introduction, data collection, data processing and analysis, the final manuscript, and supplementary material are all arranged.
+The discussion portion can be worked on a little
 
 ### Summary assessment
-
 * well structured
@@
 
-The project is documented very well.
-I was able to understand each step of the entire analysis, each decision that was made, and almost every line of code.
-There was one section of code in the main_analysis script that didn't have comments explaining what it does, but I know enough about R to figure it out (multivariate regression).
-Adding in a couple comments there would eliminate that issue.
-There is enough information provided otherwise in the code or Rmd file for documentation of all the steps.
+Very well documented
 
 ### Summary assessment
-
 * fully and well documented
@@
 
-All the results are fully reproducible without any issues.
-READMEs provide enough information for what steps to take to reproduce the analysis, and no manual intervention is needed.
-I'm able to re-run the entire analysis without editing anything.
-Nothing to improve here.
+All results were reproducible
 
 ### Summary assessment
-
 * fully reproducible without issues
@@
 
-The study wasn't incredibly thorough, not many alternatives to processing or analyzing the data were discussed.
-A regression analysis was completed with univariate and multivariate models for the first question, but no other models or processing approaches were tried.
-My suggestion would be trying a machine learning model like a LASSO regression or a tree/forest just to see if you could get a new result (RMSE, etc.) from a different perspective, instead of just moving onto the next one right away.
-But it could be that you did, and didn't include it in the final repository.
-You could also try both bottom-up and top-down subset selection in the multivariate regression analysis, that's another option.
-Only the LASSO model was used for the second question, but it addressed it thoroughly enough to answer and conclude.
-
+Questions and hypothesis were addressed. More alternatives can be considered
 ### Summary assessment
-
 * decent level of thoroughness
@@
 
-Overall, very interesting topic and great project!
-Fully reproducible from beginning to end, and successfully analyzes data from multiple sources to answer the two questions.
-I think spending a little time addressing the above comments would be beneficial to the overall condition of the project.
+_Add any other comments regarding the overall project here. Write anything you think can help your classmate improve their project.
+
+Great project considering the situation we are in. all results reproducible and no manual intervention was needed.

### 2021-12-11 products/manuscript/Manuscript.Rmd
Changes based on the peer review. Final Editing (hopefullyy)

--- 2021-11-28 products/manuscript/Manuscript.Rmd
@@
 COVID-19 has been the talk of the world for the past two years, and it is still prevalent.
-The affects of COVID-19 not only affected individual people, but it also affected businesses, health protocols, and the economy of governments.
-It has affected all countries not only specific to a type of country which is the scary part.
+COVID-19 not only affected individual people, but it also affected businesses, health protocols, and the economy of governments.
+All countries faced this impact no matter their location or how developed their country is.
 It is truly indeed a pandemic.
@@
 I am interested in looking at the difference between the proportion of those who recovered to those who have died for each country in the world.
+The variables I am considering are: percent of cases within a population, continent the country is on, number of tests per person, gdp per capita, proportion of those vaccinated, government type, and health care funds.
+By looking at some simple linear regressions and other exploratory analysis, the fit of the model to the data were not good for each simple linear regression.
+I decided to do a step-wise selection process, and found that the percent of cases and the continent the country resides in has an association to the difference in proportions.
+However, the fit is still not good at all.
+
+The higher the percent of cases there is in a country, the higher recovery proportion there is when keeping the continent constant.
+North American countries as a whole has a higher death proportion compared to African countries as a whole when keeping the percent of cases constant.
+
 I am also looking at how the proportion of people vaccinated in each country is affected by multiple factors.
@@
 
-By looking at some simple linear regressions and other exploratory analysis, I determined due to how high the recovery rate is within each country overall and no association based on R-squared values, the difference in proportions did not change based on multiple factors.
-
 While looking at vaccination proportions, I decided to do some analyses using training/test data sets and applying them to find the RMSE of the full model and a null model.
 When finding the RMSE, I found that the full model does a better job in fitting the data compared to the null model.
-This is a good sign to share that some combination of predictors have an association to the proportion of vaccinated people for each country.
+This is a good sign in showing that some combination of predictors have an association to the proportion of vaccinated people for each country.
 
-The next analysis was creating a machine learning modeling technique called the LASSO model.
+The next analysis was using a machine learning modeling technique called the LASSO model.
 This type of modeling will create multiple models from a five-fold cross-validation and a tuning grid which will try to find the best performing model.
-From the LASSO we find that model 25 give the best model.
-
-**Need to find a way to extract this model**
+From the LASSO, we find that model 27 gives the best model, and the percentage of cases in a country, which continent the country resides in, and the type of government have impacts to the vaccination proportions.
+Compared to Africa, South America and Asia increases the proportion of vaccinations the most while an increase of cases means more vaccinations as well.
+A monarchy government compared to a republic government increases the proportion of people vaccinated.
 
@@
 COVID-19 hit the world by storm and is a possible ever-lasting change to the way society acts.
-It also had an impact on each countries funding towards health care.
+It also had an impact on each country's funding towards health care.
 Actions within the past two years have been taken to help fight against this virus.
-Although COVID-19 has been here for over two year, there is still problems with under staffing in hospitals and not enough beds at some hospitals.
+Although COVID-19 has been here for over two years, there are still problems with under staffing in hospitals and not enough beds at some hospitals.
 Recovery rates are still high in the United States but what about the whole world.
 Other countries are still getting hit hard by this virus.
-In terms of vaccination output, there has been recent back and forth movements on people getting the vaccination.
-Although in developed countries this happens but what about the less developed countries and regions such as Africa.
+In terms of vaccination output, there has been back and forth movements on people getting the vaccination.
+A poll, close to when vaccinations were starting to disperse, showed that only 51% of adults in the United States said they would get a vaccine [@choi21].
+Although this happens in developed countries, what about the less developed countries and regions such as Africa.
 While people in the United States are fighting about whether the vaccine should be given or not, the people in these other countries might have a limited supply.
+According to the United Nations in September of 2021, there is still an unequal amount of vaccinations going to developed and less developed countries.
+This could be due to the cost of each vaccination [@UN21].
+In a paper by Choi, as of January 2021, cases were fairly even between developed and less developed countries [@choi21].
+This just shows that although COVID-19 is affected everyone, the supply is not equal between all countries.
+In a paper by Mathieu et al., they show a world map and the distribution of vaccines by country as of April of 2021.
+We can see that Africa is the continent that has the lowest distribution of vaccines [@mathieu21].
 
-Looking at the difference in recovered patients and those who have died will be beneficial in determining if more care is needed in certain countries.
-I also think the percentage of people vaccinated is also affected and can lead to important results that might show we need more either distribution or requirements of the vaccines.
+One of the goals of this analysis is to look at the difference in proportions of recovered patients and those who have died.
+This will be beneficial in determining if more care and funding is needed for countries.
+According to  paper by Liang et al., the mortality rates were negatively impacted by the number of tests performed and also based on the countries economic status and government [@liang20].
+
+My other goal for this project is to analyze how the percentage of people vaccinated is affected.
+It could possibly lead to important results that might show that we need more either distribution or requirements of the vaccines.
+As mentioned above, there is a uneven distribution of vaccinations but is this significantly different and what other factors could affect vaccination numbers?
 
@@
 These data sets includes total covid-19 cases, active cases, total deaths, total recovered, total tests, and population size.
-This data was updated on kaggle as of Sept 16th.
+This data was updated on kaggle as of September 16, 2021.
 Another data set I am using is a vaccination data set for each country.
-This is also data as of Sept.
-16th.
-The other data sets are providing single stats for each country that was not included in the larger data sets but I still wanted to use.
+This is also data as of September 16, 2021.
+The other data sets are providing statistics for each country that was not included in the larger data set but I still wanted to use.
 These data sets include the following statistics/demographics: healthcare funding for 2019-2020, vaccination numbers/percentages, gdp per capita, and government type.
 
-I plan on combining the data sets into one with the variable combining them all is the country.
+I combined the data sets into one using the countries name as the merging variable.
 If there are missing data within the variables used in analysis, I will exclude those countries.
-There will be a lot of cleaning for this data just to make it as I want it and work.
+There will be a lot of cleaning for this data just to make it as I want it.
+
+The variables I will be using in my analysis are:
+
+* **prop_diff_recov_vs_death** - proportion of those who have recovered versus those who have died from COVID-19 (for those who has gotten COVID-19)
+* **Value** - health care funds (in U.S. dollars)
+* **prop_vacc** - proportion of those who are vaccinated (for the whole population of a country) 
+* **pct_cases** - percent of the population in each country that has gotten COVID-19
+* **location** - the continent in which the country resides in 
+* **test_per_person** - the number of tests a person has taken (on average)
+* **gdp_per_capita** - the gdp per capita for each country (in U.S. dollars)
+* **prop_death** - proportion of those who have died from COVID-19 
+* **government** - Type of government (Republic, Monarchy, or Provisional)
 
@@
 ## Questions/Hypotheses to be addressed
-1) Since this data is cumulative and does not break down the amount of cases per month per country, I want to compare the recovery proportion to the death proportion and if this difference is based on variables such as the number of vaccines pushed out to society, the amount of funding the health care system has, the amount of tests taken.
+1) Since this data is cumulative and does not break down the amount of cases per month per country, I want to compare the recovery proportion to the death proportion.
+I want to see if this difference is based on variables such as the proportion of those vaccinated, the amount of funding the health care system has, the amount of tests taken per person, the location of the country, the percent of cases in a country, the gdp per capita, and the government type.
 This will hopefully show if the amount of health care support really helps patients recover.
 
-2) I will also look at how the percent of people vaccinated for each country is affected by multiple predictors.
+2) I will also look at how the percent of people vaccinated for each country is affected by multiple predictors (location of country, tests per person, government type, proportion of deaths, gdp per capita, and percent of cases per country).
 I am planning on seeing that more developed countries based on GDP per capita will have a more vaccinated population.
@@
 Some confounders that I could run into are:
+
 1) Developed countries might have more cases as more travelers visit these countries and it could have an effect on recovery numbers.
@@
 ## Data import and cleaning
-For data importing and cleaning, I first had to import ten different data sets into R from multiple sources specified in the Description of data and data source.
-After this importation, The first thing to do before combining the data set was to create indicator variables for each country.
+For data importing and cleaning, I first had to import ten different data sets into R from multiple sources specified in the description of data and data source section.
+After this importation, the first thing to do before combining the data set was to create indicator variables for each country.
 After this, I wanted to make sure that each variable was the right class as some of the numeric variables were character variables.
@@
 I then was able to combine each continent data set into one world data set to make the importing go easier.
-After combining the continent data sets, I need to only include the variables from the world data set needed for analysis and make sure that all missing values were removed as I only want full information observations.
+After combining the continent data sets, I need to only include the variables from the world data set needed for analysis and make sure that all missing values were removed as I only want full observations.
 However this would delete some countries that were large in population such as China and the Philippines as they did not have vaccine information.
@@
 However, the country names for all datasets were different.
-One included a three letter code for countries, then another ones used different abbreviations/names for countries.
-I had to use the countrynames package to help change into names I wanted for the three-letter code.
+One included a three letter code for countries and another one used different abbreviations/names for countries.
+I had to use the 'countrynames' package to help change into names I wanted for the three-letter code.
 Since I could not change all of the countries names to the way I wanted it, I had to unfortunately go into the raw data csv files to change them manually for some of the continent specific data sets.
@@
 I was able to do these in r because there were not as many country names to change compared to the other data sets I had to manually change.
-A comment within the proccessingscript.R tell of the countries that needed manual changes.
+A comment within the proccessingscript.R tells of the countries that needed manual changes.
 After all of this was good to go, I used the merge function to combine all of the data sets the way I wanted to.
 
-I created some variables from the pre-existing ones that I will use within my analysis. Some of these changes were, for example, creating proportions from percentages and combining two variables.
+I created some variables from the pre-existing ones that I will use within my analysis.
+Some of these changes were, for example, creating proportions from percentages and combining two variables.
+These variables can described above in the description of data and data source section.
 
@@
 
+### Methods
+
+For question 1's analysis, I first decided to do some simple exploratory analysis on the data.
+This included some simple linear regression models just to see how each variable fits with my outcome.
+After fitting most of the predictors to the outcome, I decided to look at a step-wise procedure which will hopefully create the best model for this data.
+Before fitting the variables into step-wise selection, I did a correlation matrix to see if some variables are very correlated.
+After considering the correlation matrix, the variables I considered in my analysis were the proportion of those vaccinated, tests per person, continent the country resides in, gdp per capita, government type, and percent of cases in a country as predictors.
+
+I used the caret package within R to do this step-wise procedure.
+After doing the step-wise procedure and finding which variables will stay in the model, I created that model just to see how well the model actual is.
+I created a predicted versus observed values scatterplot and a supplementary residual plot as well.
+These plots and the model model will help give me an idea of what variables are useful in predicting the difference in proportions of recovered people and people who have died.
+
+### Exploratory Analysis
+
 Table 1.1 displays the top five and bottom five death rates.
 Also within this, it shows their corresponding recovery rates as well.
-We can see that the bottom five death rates are from Asia but also the highest, Yemen, is also within the Asia continent as well.
-<br /> 
+We can see that the bottom five death rates are from Asia but also the highest, Yemen, is within the Asia continent as well.
+Yemen has a much higher death rate compare to the other top countries.
+It is more than double the second highest.
+However, the sample size of total COVID-19 cases is smaller than others.
 <br />
@@
 
-Using the logit transformation, I ran 4 simple linear regression models. Below I provide the statistics of one of the models within Tables 1.2.
-
+Using the logit transformation, I ran 4 simple linear regression models.
+Below I provide the statistics of one of the models within Tables 1.2.
 Table 1.2 is showing statistics for the predictor vaccine percentage.
@@
 Table 1.2: Summary statistics for linear regression (logit transformation of prop diff of recovery and deaths vs. vaccine)
+
 ```{r table 1.2,  echo=FALSE}
@@
 
-Looking at Table 1.2, we can see that the R^2 values are very low.
-This is similar to all of the other simple linear regression models.
+Looking at Table 1.2, we can see that the R^2 values are very low, and the deviance is also somewhat high.
+This is a similar situation to all of the other simple linear regression models.
 The statistics for the other three models are within the supplemental material.
-Therefore none of these predictors shows any association with the logit transformation outcome.
-Since not even one predictor is good for this outcome, I am going to stop my analysis with this outcome and go to my second question which uses the percent of cases within each country.
+Therefore none of these models are good fits to the data.
+I am going to see if a step-wise selection of predictors could increase the fit of the model to the data.
+
+### Main Analysis
+
+The main analysis pertains to using step-wise selection for the variables I specified within the methods section for question 1.
+I am doing step-wise selection because for each simple linear regression, the r-squared values are very small and I wonder how my model will do if I try to select the best variables in predicting the logit transformation of the difference in proportions between recovery and deaths.
+
+Table 1.3 shows the resulting model from the step-wise selection.
+If you want more details on this selection, there is a table showing the process of model selection within the supplementary material.
+We can see that the two variable that were selected using step-wise selection were percent of cases in a country and the location or continent the country is in.
+Although these two variables were selected and they are significant on at least one level, we can see that the fit according to R-squared is still very low.
+
+Table 1.3
+```{r table 1.3,  echo=FALSE}
+tab13=readRDS("../../results/summ_stepwise_tab.rds")
+knitr::kable(tab13)
+
+tab132=readRDS("../../results/stepwise_stat_tab.rds")
+knitr::kable(tab132)
+```
+
+This low fit can be shown within Figure 1.3 and Figure 1.4 where they show the predicted values versus observed values as well as the corresponding residual plot.
+The data points are spread out and it seems, within the residual plot, it shows a funnel shaped plot which means heteroskedasticity.
+Therefore, possibly more work could be done to this data/question to get a better fit such as a transformation on the outcome.
+However, the outcome is already transformed with a logit transformation.
+
+```{r figure 1.3, echo=FALSE}
+knitr::include_graphics("../../results/pred_vs_obs_stepwise.png")
+```
+
+```{r figure 1.4, echo=FALSE}
+knitr::include_graphics("../../results/resid_stepwise.png")
+```
 
@@
 
-For my main analysis, I will be using the logit transformation of the proportion of people vaccinated within each country as my main outcome.
-Since I am using the logit transformation, I am creating another data set that is a subset of the overall data set that only includes the variables I will be using for my analysis.
+For my analysis on question 2, I will be using the logit transformation of the proportion of people vaccinated within each country as my main outcome.
+Since I am using the logit transformation, I am creating another data set that is a subset of the overall data set which only includes the variables I will be using for my analysis.
 The predictor variables I will be testing out on the main outcome are: percent of cases within a population, continent the country is on, number of tests per person, gdp per capita, proportion of those who died, government type, and health care funds.
@@
 
-After comparing the two RMSE, I will be going into more detail on model fitting and tuning.
-I will be applying my full model a machine learning model, the LASSO model.
-This will help find the best model I can use for the predictors in predicting the logit transformation of the proportion of people vaccinated.
-I will be applying this lasso model to both my test and training data created before to make sure the model fits both data sets similarly so show improvement on over-fitting.
+After comparing the two RMSE's, I will be going into more detail on model fitting and tuning.
+I will be applying my full model to a machine learning technique, the LASSO model.
+LASSO stands for the Least Absolute Shrinkage and Selector Operator.
+The reason why I am choosing the LASSO model compared other machine learning models such as a simple tree or a random forest is because LASSO is in the middle of simplicity and complexity which is what I describe the simple tree and the random forest, respectively.
+Using the LASSO model will try to find the best performing model for the data.
+This uses a tuning process and a 5-fold cross validation.
+This process will penalize variables if it is not useful in predicting the outcome.
+This penalty will not remove the variables but instead penalize these variables to values of zero or close to zero where they wont have a large impact on the model.
+This will help find the best model to predict the logit transformation of the proportion of people vaccinated.
+I will be applying this lasso model to both my test and training data to make sure the model fits both data sets similarly so show improvement on over-fitting.
 
@@
 
-Table 2.1 displays percentage of people vaccinated within the top five and bottom five countries.
+Table 2.1 displays the percentage of people vaccinated within the top five and bottom five countries.
 Just by the look of the countries displayed, the less developed countries tend to not have a lot of people vaccinated.
@@
 
-When looking at Figure 2.2, we can see that those countries who have large values of GDP per capita, their population is more vaccinated.
-The vaccination number varies when the GDP per capita is relatively low.
+When looking at Figure 2.2, we can see that those countries who have large values of GDP per capita tend to have their population more vaccinated.
+The vaccination proportion varies when the GDP per capita is relatively low.
 
@@
 
-Looking at Figure 2.3, this side-by-side boxplot shows that Africa is significantly different to all of the other continents.
-Africa is historically known is a less developed continent.
+Looking at Figure 2.3, this side-by-side boxplots show that Africa is different to all of the other continents.
+Africa is historically known as a less developed continent.
 This could be result in lower supplies to Africa compared to other places in the world.
@@
 Figure 2.4 is showing the logit transformation of the proportion of the population vaccinated.
-This is due to the skewness of the non-transformed vaccinatio numbers.
+This is due to the skewness of the non-transformed vaccination numbers.
 We see that this is still skewed but its skewed left where most of the data is around that 0 value.
@@
 
-As specified within the methods section, the first part of my analysis will be comparing the RMSE of my full model to the null model.
+As specified within the methods section of question 2, the first part of my analysis will be comparing the RMSE of my full model to the null model.
 However, before I do this I will run a correlation matrix of the numeric variables to see if some variables are highly correlated.
@@
 
-Table 2.2: Correlation martrix of predictors and outcome
+Now lets see how our outcome, prop_vacc, is correlated with our predictors.
+Proportion of vaccinated people has roughly an 0.57 correlation with both Value which is health care funds and gdp per capita.
+This was something I was curious about, if countries with a lot of funding could have a direct association to vaccinations.
+We will see in our main analysis to see if gdp per capita is significant in predicting the proportion of vaccinated people.
+
+Table 2.2: Correlation matrix of predictors and outcome
 ```{r table 2.2,  echo=FALSE}
@@
 
-Now that I have my full model solidified, Table 2.3 shows the summary table and some other statistics of this full model.
+Now that I have my full model solidified, Table 2.3 shows the summary table and some other statistics of this full model where the outcome is the logit transformation for the proportion of vaccinated people.
+
 We can see that our adjusted R-sqaured value is around 0.64 and most of our factors are significant except the proportion of this who died and the COVID tests per person.
@@
 Once we have our model type and our recipe, we can combine them and put it in a workflow.
-This workflow will be able to combine the two `settings` we have and create a fit object which eventually creates the statistics of our models and our predictors.
+This workflow will be able to combine the two 'settings' we have and create a fit object which eventually creates the statistics of our models and our predictors.
 However, since our main objective is looking at the fit of our data, we want to look at the RMSE.
@@
 Table 2.4 shows the two values for the full model where the top is the training value and the bottom is the testing value.
-We can see that these two values are somewhat similar and possibly we want to a bit closer.
+We can see that these two values are somewhat similar and possibly we want it to be a bit closer.
 However, these two values are relatively low which is a good sign.
 
-Comparing Table 2.4 to Table 2.5, which shows the null models results, we can see our full model does a lot better than the null model.
+Table 2.5 has the same setup as Table 2.4 where the top value is the training rmse and the bottom is the testing rmse.
+We can see that these two values are very high but they are somewhat close to each other but not as close as the full model rmse's.
+
+Comparing the two models, we can see our full model does a lot better than the null model.
 The RMSE is a lot worse for the null model.
 This is a good sign for the data as we now know that these predictors do have some type of association since it is way better than the null model.
-We can now proceed to creating our LASSO model and tuning this full model to find the best performing model.
+We can now proceed to creating our LASSO model and tuning it to find the best performing model.
 
@@
 This will produce the variety of models.
-To find the best model, we an apply a function called show_best and specify RMSE to find the best model in terms of RMSE.
+To find the best model, we an apply a function called 'show_best' and specify RMSE to find the best model in terms of RMSE.
 After putting this model into a workflow and creating a fit object, we get the statistics of the best model.
@@
 We can see this more within Figure 2.6 which is the residual plot.
-Within Figure 2.6, we could see a slight pattern at the begging points where they are grouped up more.
+Within Figure 2.6, we could see a slight grouping of the points around the predicted value of -2.5.
 
@@
 The last part of the LASSO modeling analysis is fitting the model to the test data to see how well they both look together.
-To look at the test data results, we need to run a last fit function which just applies the test data to the best model found above.
+To look at the test data results, we need to run a 'last fit' function which applies the test data to the best model found above.
 Figure 2.7 and Figure 2.8 show the predicted values versus observed and its residual plot for the test data.
-Just like the training test, most of the predicted values do have a similar trend and it close to the observed value.
-The residual plot shows that concerning line of observations for the lower values.
-This residual plot show more variability in the lower values more.
-Table 2.6 shows the RMSE values created from this LASSO model.
-We can see unfortunately these are overall higher compared to the full model numbers.
-It is still better than the null model however.
+Just like the training data, most of the predicted values do have a similar trend and it close to the observed value.
+The residual plot shows that grouping of observations for the lower values.
+This residual plot show more variability in the lower values as well.
 
@@
 
+Table 2.6 shows the RMSE values created from this LASSO model.
+We can see unfortunately these are overall higher compared to the full model numbers which were around 1.08 and 1.16 for the training and test data, respectively.
+It is still better than the null model however (1.85 and 1.95).
+
 Table 2.6: Comparing RMSE for training and test split data for LASSO model, respectively 
 ```{r table 2.6,  echo=FALSE}
-tab26=readRDS("../../results/LASSO_rmse_train_tab.rds")
+tab26=readRDS("../../results/LASSO_rmse_train_test.rds")
 knitr::kable(tab26)
-
-tab262=readRDS("../../results/LASSO_rmse_test_tab.rds")
-knitr::kable(tab262)
 
@@
 However these estimates are based on the logit transformation of our outcome.
-The last column is the transformation to display the transformed of each predictor applying the exp() function to the estimates.
+The last column is the transformation from the logit value to help in more practical predictions.
 
-From the estimates, we can see although there was 2 variables that suppressed to a value of 0 based on the LASSO model, the predictor gdp per capita have roughly a value of 0 which translates to a odds ratio close to 1.
-Those who include 1 or close to 1 results in a non-significant variable in predicting the outcome.
+From the estimates, we can see although there was 2 variables that suppressed to a value of 0 based on the LASSO model, the predictor gdp per capita have roughly a value of 0.
+Therefore the LASSO model did surpress this variable as well.
+It just was not exactly to 0.
 We can see the continent that the country is in also has an affect on proportion vaccinated where the continent Africa is the reference group.
-The continents that increases the proportion of being vaccinated the most are Asia and South America.
-We can also see that in terms of government type where a republic government is the reference group, those countries that are a monarchy increase the proportion of having more vaccinations.
-In opposite affect, a provisional government country lowers the proportion of getting people vaccinated.
+The continents that increase the odds for a higher proportion of being vaccinated the most are Asia and South America.
+We can also see that in terms of government type where a republic government is the reference group, those countries that are a monarchy increase the odds of a higher proportion of vaccinations.
+In opposite affect, a provisional government country lowers the odds of getting more people vaccinated proportionally.
 
@@
 The analysis was split up into two different analyses specified by the two questions.
+
 The first analysis was looking at the difference of proportions between recovery and death rates.
-With some exploratory analysis, no association according to the simple linear regression between multiple predictors and the outcome.
-Since within some exploratory analysis and very small r-square values, I decided to stop that analysis on the first question and go to the next question.
+With some exploratory analysis, there was low fitting according to the simple linear regression between multiple predictors and the outcome.
+Due to the very small r-squared values, I decided to do a step-wise selection on the predictors.
+In result, I found that percent of cases and the continent is the remaining two variables.
+The only level from the location variable that is significantly different from the reference group, Africa, is North America.
+Although they show a significant impact in predicting the logit transformation of difference in proportions, we can see that step-wise selection does not create a model that fits the data well.
+While looking at the residual plot, we see a funnel shaped plot which means heteroscedasticity and possibly more transformation options could be applied to make this model fit better.
+However in this current state, the variables do not fit the data well.
+If we look at the estimates and transform them out of the logit transformation, we find that if we increase the percentage of cases within a country, there will be a slight increase in odds of having a higher recovery proportion compared to the death proportion with the location staying constant.
+Contrary, North America countries, compared to African countries as a whole, decreases the odds of having a higher difference in proportions between recovery and death.
 
-<\ br>
-
-The second question pertains to vaccination proportions.
+The second question pertains to how vaccination proportions are affected by multiple predictors.
 Using the logit transformation and applying the LASSO modeling, out of the 6 predictors within the full model, only 4 remained within the equation.
 These include percent of cases for a country, gdp per capita, the continent the country resides, and the type of government.
-Out of these, those who have the most impact are location and government type, and gdp per capita has roughly an estimate of 0 which probably has no impact on the proportion of people vaccinated.
+Out of these, those who have the most impact are location and government type, and gdp per capita.
+However, gdp per capita has roughly an estimate of 0 which probably has no impact on the proportion of people vaccinated.
+With the reference for the location variable being Africa, we can see that no matter which level of location you look at, they increase the odds of having a higher proportion of people getting vaccinated while keeping all other variables constant.
+The type of government being significantly different could be due to just the small amount of countries that classify as monarchy and provisional compared to republic or it could really be that significantly different.
+Regardless, the monarchy type of government increase the odds that the proportion of vaccinated people is higher compared to a republic type of government, and a provisional government decreases the odds that the proportion of vaccinated people is higher compared to the republic government given all other variables constant.
 
 ## Strengths and Limitations
-Some strengths to this analysis was that since this topic of COVID-19 is in todays news which means much of the countries information is available in public use.
-Along with COIVD-19 information, most information of countries are of public use therefore lots of information  was easy to find.
+Some strengths to this analysis was that since this topic of COVID-19 is in today's news, a lot of each country's information is available in public use.
+Along with COIVD-19 information, most information of countries are of public use.
+Therefore a lot of information  was easy to find.
 
-Some limitations for this analysis was that I was using cumulative data which means as more days pass, this data will become old analyses.
+Some limitations for this analysis was that I was using cumulative data which means as more days pass, this analysis will become old.
 Another limitation was although I got the information I wanted, I had to use multiple sources to find them which results in combining data sets and ultimately errors.
@@
 ## Conclusions
-We find that due to high high recovery right is for most countries compared to death rates, and no predictors have a large association, according to the simple linear regression r-squared values, the difference in proportion between recovery and death rates is not affected by most predictors.
+We found that the percent of cases and the continent are significantly associated with the difference in proportion of recovered and deaths.
+The higher the percent of cases there is in a country, the higher recovery proportion there is when keeping the continent constant.
+North American countries as a whole has a higher death proportion compared to African countries as a whole when keeping the percent of cases constant.
+I was curious on if the amount of health care support helps patients recover.
+However based on these results, we see that the amount of tests per person and gdp per capita does not really affect recovery rates unfortunately.
+However, the fit from my resulting step-wise selected model is still not good.
+This is constant throughout the simple linear regressions and the step-wise model.
+A different transformation could possibly help better the fit more as the residual plot shows a funnel shaped plot.
+
+If we revisit the Liang et al.
+paper, they mentioned in their analysis that the tests per person, type of government, and economic status negatively impacts mortality rates [@liang20].
+We see that within my analysis of the difference between recovery and death proportions, none of these variables had an impact on the this difference in proportions.
+This is not a direct comparison between analysis, and their analysis was during 2020.
 
 We see that the proportion of people vaccinated is associated with the amount of COVID-19 cases, which continent the country is in, and type of government.
-Given all other variable constant, Asia and South America gives the highest proportional increase compared to Africa, and monarchy governments increase vaccinations compared to republic government countries.
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+Given all other variable constant, Asia and South America gives the highest proportional increase compared to Africa.
+Compare to what Mathieu et al.
+said, as I previously mentioned above in the general background section, we do see that Africa has a significant different proportion of people vaccinated and in the negative direction.
+Monarchy governments increase vaccinations compared to republic government countries but provisional governments decrease the proportion of those vaccinated.
+One prediction I had within the questions/hypotheses section above was that countries with more gdp per capita will have higher vaccinated proportions.
+However, based on my LASSO model, gdp per captia was not significant in predicting vaccine proportions.
+If we look at more previous studies I mentioned above in the background section, the United Nations mentioned that the distribution of vaccines could be based on the actual prices of the vaccines.
+This could have a direct correlation to the gdp per capita for each country.
+However within my analysis, gdp per capita does not have much of an impact on vaccination proportions when all other variables stay constant [@UN21].
 

### 2021-12-11 products/manuscript/Supplemental Material.Rmd
Changes based on the peer review. Final Editing (hopefullyy)

@@
 
+This table shows the correlation matrix used to determine which variables I will use for the step-wise selection in question 1.
+We see that value which is the health care funds and the gdp per capita are very correlated.
+Since health care funds has a lot of missing, I decided to use gdp per capita.
+
+```{r table stepwise corr,  echo=FALSE}
+tab55=readRDS("../../results/supp_mat/cor_tab_prop_diff.rds")
+knitr::kable(tab55)
+```
+
+This table is showing the process of the stepwise selection for the logit transformed difference in proportions.
+The nvmax means the number of predictors used in each model and each row shows the best model from each quantity of variables used.
+Although the number of variable shows 12, it is only 6 distinct variables.
+I used the package caret and the function I used considers each level of nominal variables as a separate variable.
+This mean the location variable and the government variable is split up by the levels.
+We can see that nvmax value 3 has the lowest RMSE value which mean it is the best model to use.
+We can also see that the R-squared values are all very low, which is somethin I addressed in the manuscript.
+
+```{r table stepwise process,  echo=FALSE}
+tab6=readRDS("../../results/supp_mat/step_proc_tab.rds")
+knitr::kable(tab6)
+```
+
 **Supplemental material for question 2**

### 2021-12-11 code/analysis_code/analysisscript.R

@@ -11,7 +11,7 @@ library(here) #for data loading/saving
 library(tidyverse)
 library(AER)
 library(tidymodels)
-
+library(caret)
 
 
 #path to data
@@ -89,7 +89,7 @@ vacc_vs_hc = mydata %>% ggplot(aes(x=Value, y=`% of population fully vaccinated`
 #number of cases.
 prop_diff_recov_died = mydata %>% ggplot(aes(x = prop_diff_recov_vs_death )) + geom_histogram(binwidth = 0.05) + 
   ggtitle("Figure 1.1: Histogram of Difference in proportions \n of those who recovered versus those who died \n (by Country)") + 
-  xlab("Proportion difference between recovery rates and death rates")
+  xlab("Proportion difference between recovery rates and death rates") + theme(plot.title = element_text(hjust=0.5))
 
 
 #Looking at the distribution using a histogram
@@ -166,7 +166,7 @@ mydata$logit_prop_diff = logit(mydata$prop_diff_recov_vs_death)
 #number of cases.
 log_prop_diff_hist = mydata %>% ggplot(aes(x = logit_prop_diff )) + geom_histogram(binwidth = 0.5) + 
   ggtitle("Figure 1.2: Histogram of logit transformation of difference \n in proportions of those who recovered versus those who died \n (by Country)") + 
-  xlab("Logit transformation of the proportional difference between recovory and death rates")
+  xlab("Logit transformation of the proportional difference between recovory and death rates") + theme(plot.title = element_text(hjust=0.5))
 log_prop_diff_hist
 
 #Saving figure * to a location for future references
@@ -220,13 +220,108 @@ tabloc1_file = here("results/supp_mat", "tableloc1.rds")
 saveRDS(loc_stats, file = tabloc1_file)
 
 
+
 #After looking at the models, it looks like the difference between the proportion of recovery and deaths
-#are not affected by the predictor variables. After finding these results and looking at the data, the problem might live within
-#the fact that over half of the difference in proportions are over 0.9. 
+#are not affected by the predictor variables individually. After finding these results and looking at the data, the problem might live within
+#the fact that over half of the difference in proportions are over 0.9. However, I will try a subset selection of manually selecting the predictors
+#and see if a variable become significant.
 
-#Therefore, my next outcome I will look at will be just the case percentage for each country.
[…]

### 2021-12-11 code/analysis_code/main_analysis.R

@@ -42,20 +42,23 @@ saveRDS(cor_tab, file = cor_file)
 
 #Visualization of the simple linear regression models
 
-#This is a table showing the top 5 and bottom 5 percentage of vaccinated people 
+#This is a table showing the top 5 and bottom 5 percentage of vaccinated people. I first arranged the countries by the perfect on population vaccinated
+#Then I used the head and tail functions to display the top 5 and bottom 5 countries
 vacc_tab = mydata %>% select(Country, location, `% of population vaccinated`) %>% 
   arrange(desc(`% of population vaccinated`)) %>% {rbind(head(., 5), tail(., 5))}
 vacc_file = here("results", "vacc_tab.rds")
 saveRDS(vacc_tab, file = vacc_file)
 
-#This is a histogram showing the distribution of those who are vaccinated. We can see that it is somewhat uniform
+#This is a histogram showing the distribution of those who are vaccinated. We can see that it is somewhat right skewed.
+#I added the bin width just to show more of the distribution.
 vacc_hist = mydata %>% ggplot(aes(x = `% of population vaccinated`)) + geom_histogram(binwidth = 5) + 
-  ggtitle("Figure 2.1: Histogram of Percent of People Vaccinated \n (by Country)")
+  ggtitle("Figure 2.1: Histogram of Percent of People Vaccinated \n (by Country)") + theme(plot.title = element_text(hjust=0.5))
 figvacc_file = here("results", "vacc_hist.png")
 ggsave(filename = figvacc_file, plot= vacc_hist)
 
 
-#Scatterplots of these simple linear regressions
+#Scatterplots of these simple linear regressions - For proportion vaccinated versus percent of population as cases
+#, gdp per capita, and tests per person
 case_scatterplot <- mydata %>% ggplot(aes(y=prop_vacc, x = pct_cases)) + geom_point() + 
   ggtitle("Scatterplot of percentage of cases and \n proportion of people vaccinated \n (by Country)") 
 fig10_file = here("results/supp_mat","cases_vs_vacc.png")
@@ -63,13 +66,14 @@ ggsave(filename = fig10_file, plot=case_scatterplot)
 
 gdp_per_cap_scatterplot <- mydata %>% ggplot(aes(x=gdp_per_capita, y = prop_vacc)) + geom_point() + 
[…]

### 2021-12-11 code/processing_code/processingscript.R

@@ -97,6 +97,8 @@ clean_funds2$Country = countrycode(clean_funds2$LOCATION, "iso3c", "country.name
 clean_funds2$Country = ifelse(clean_funds2$Country == "Czechia", "Czech Republic", 
                               ifelse(clean_funds2$Country == "United Kingdom", "UK", clean_funds2$Country))
 
+#This is changing some names of countries to make sure they all combine properly between data sets
+
 gdp <- gdp %>% mutate(ifelse(Country == "United Kingdom", "UK", 
                              ifelse(Country == "Bahamas, The", "Bahamas", 
                                     ifelse(Country == "Brunei Darussalam", "Brunei", 
@@ -113,7 +115,7 @@ gdp <- gdp %>% mutate(ifelse(Country == "United Kingdom", "UK",
                                                                                                                  ifelse(Country == "Korea, Rep.", "South Korea", Country))))))))))))))) %>%
   mutate(Country = `ifelse(...)`) %>% select(-`ifelse(...)`)
 
-#Cleaning the vaccine data set
+#Cleaning the vaccine data set by removing all of the missing values
 
 clean_vaccine = drop_na(vaccine)
 
